---
title: "Proxy $MSY$ Reference Points for data poor stocks"
subtitle: "Power of length based Indicators to detect recovery"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
       includes:
            in_header: header.tex 
---

## Summary
+ The OM has been rerun where F after recovery is 0.7$F_{MSY}$. 
+ The reference period is 61:120 and the period of overfishing is 80:109. this gives 60 years for which 30 are overfished.
+ For overfishing reference period is 61:100, with overfishing occuring in 80:100
+ For recovery reference period is 100:120, with recovery occuring in 110:120

\newpage

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE,
               fig.width =10, 
               fig.height=12,
               cache     =TRUE, 
               fig.path  ="../tex/roc-final-",
               cache.path="../cache/roc-final/",
               dev       =c("png"))

iFig=0
iTab=0
```

\newpage
```{r, pkgs} 
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)
library(mydas)
library(ggplotFL)

library(plyr)
library(dplyr)
library(reshape)
library(GGally)

library(popbio)

library(LBSPR)
library(spatstat)
```

```{r, params, eval=FALSE}
## Read in parameters from fishbase
load(url("https://github.com//fishnets//fishnets//blob//master//data//fishbase-web//fishbase-web.RData?raw=True"))

## lh
lh=subset(fb,species%in%c("Pollachius pollachius","Psetta maxima",
                          "Scophthalmus rhombus","Raja clavata",
                          "Sprattus sprattus","Sprattus sprattus sprattus")) 

spp=c("Pollachius pollachius","Psetta maxima",
      "Scophthalmus rhombus","Raja clavata",
      "Sprattus sprattus","Sprattus sprattus sprattus")
nm =c("Pollack","Turbot","Brill","Ray","Sprat","Sprat")

names(lh)[c(14:17)]=c("l50","l50min","l50max","a50")
lh=lh[,c("species","linf","k","t0","a","b","a50","l50","l50min","l50max")]

lh[is.na(lh["l50"]),"l50"]=(lh[is.na(lh["l50"]),"l50min"]+lh[is.na(lh["l50"]),"l50max"])/2
lh=lh[,-(9:10)]

# Add extra parameters for pollack
lh[lh$t0>=0,"t0"]=NA
lh=rbind(lh[1,],lh)
lh[1,-c(1,7:8)]=c(84.6,0.19,-0.94,0.01017,2.98)
lh[1,"l50"]=0.72*lh[1,"linf"]^0.93

lh=transform(lh,l50linf=l50/linf)

fctr=ddply(lh,.(species),with,mean(linf))
fctr=as.character(fctr[order(fctr$V1),"species"])

lh$species=factor(as.character(lh$species),levels=fctr)
```

```{r, lhs, eval=FALSE}
lhs=dlply(lh,.(species), with, {
  
    res=lhPar(FLPar(
                   linf =mean(linf,na.rm=T),
                   k    =mean(k,   na.rm=T),
                   t0   =mean(t0,  na.rm=T),
                   l50  =mean(l50, na.rm=T),
                   a    =mean(a,   na.rm=T),
                   b    =mean(b,   na.rm=T)))
          
     res=rbind(res,FLPar(lopt =2/3*mean(res["linf"])))
     # also Linf*(3/(3+M/K)) 
     res=rbind(res,FLPar(lmega=1.1*mean(res["lopt"])))
     res})

cor=cor(model.frame(lh)[,c("linf","k","l50","t0","a","b","l50linf")],
        use="pairwise.complete.obs")

save(lh,lhs,cor,file="data/lhs.RData")
```

```{r, design}
load("/home/laurence-kell/Desktop/projects/mydas/papers/roc/data/lhs.RData")

## Scenarios
design=data.frame(s      =c( 0.9, 0.7, 0.9, 0.9, 0.9, 0.9),
                  sel2   =c( 1.0, 1.0,5000, 1.0, 1.0, 1.0),
                  sel3   =c(5000,5000,5000,  50,5000,5000),
                  nsample=c( 500, 500, 500, 500, 250, 500),
                  m      =c(rep("gislason",5),"constant"))
design=mdply(expand.grid(Stock=names(lhs),CV=c("0.3","0.5","AR"),stringsAsFactors=FALSE), 
             function(Stock,CV) design)

key=rep(c("Base","h=0.7","Flat","Dome","Sample Size","M"),15)

save(design,key,file="/home/laurence-kell/Desktop/papers/roc/results/design.RData")
```

```{r, cor, fig.height=10,fig.width=10}
load("/home/laurence-kell/Desktop/projects/mydas/papers/roc/data/lhs.RData")

my_smooth <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
    geom_point(...,size=.5)+
    geom_smooth(...,method="lm",se=FALSE)}

my_density <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
    geom_histogram(...,lwd=1)}

theme_set(theme_bw(base_size=20))

spp=c("Pollachius pollachius","Psetta maxima",
      "Scophthalmus rhombus","Raja clavata",
      "Sprattus sprattus","Sprattus sprattus sprattus")
nm =c("Pollack","Turbot","Brill","Ray","Sprat","Sprat")

names(nm)=spp

dat=transform(lh[,c(1:3,8,6)],linf=log(linf),k=log(k),l50=log(l50))
dat=transform(dat,name=nm[as.character(species)])[,-1]
dat$name=factor(dat$name,levels=c("Sprat","Brill","Turbot","Pollack","Ray"))
ggpairs(dat,
          mapping = ggplot2::aes(colour=name,fill=name),
          lower = list(continuous = wrap(my_smooth)),
          diag=list(continuous=wrap(my_density,alpha=1)),
          title = "")+
  theme(legend.position ="none",
        panel.grid.major =element_blank(),
        axis.ticks       =element_blank(),
        axis.text.x      =element_blank(),
        axis.text.y      =element_blank(),
        panel.border     =element_rect(linetype = 1, colour="black", fill=NA),
        text = element_text(size = 30))
```

**Figure `r iFig=iFig+1; iFig`** Life history parameters and the correlations between them.


```{r, oms, eval=FALSE}
f=FLQuant(c(rep(0.1,60),seq(0.1,2.5,length.out=40)[-40],
            seq(2.5,0.7,length.out=11),rep(0.7,20)))

nits=100
set.seed(234)
srDev=rlnoise(nits,f%=%0,0.3,0)

oms=llply(lhs, function(x,f,srDev){
  eq      =lhEql(x)
  fbar(eq)=f%*%refpts(eq)["msy","harvest"]
  
  ##check fmsy is estimated
  if (any(is.na(refpts(eq)["msy","harvest"])))
    fbar(eq)[,,,,,is.na(refpts(eq)["msy","harvest"])]=
      f%*%refpts(eq)["f0.1","harvest",is.na(refpts(eq)["msy","harvest"])]
  
  om      =propagate(as(eq,"FLStock"),dim(srDev)[6])
  om      =fwd(om,fbar=fbar(eq)[,-1],
               sr=eq,residuals=srDev)
  om},f=f,srDev=srDev)

oms=FLStocks(oms)

msy=ldply(lhs, function(x){
  eq      =lhEql(x)
  model.frame(refpts(eq)["msy",1:4])})

msy=transform(msy,qname=factor(quant,labels=c("Rec","SSB","Catch","F"),
                                      levels=c("rec","ssb","yield","harvest")))
names(msy)[1]="stock"

save(oms,msy,  
     file="data/oms.RData")
```


\newpage
\blandscape

```{r, ts, fig.height=7, fig.width=12}
library(ggh4x)

load("/home/laurence-kell/Desktop/projects/mydas/papers/roc/data/oms.RData")  

p=plot(FLStocks(llply(oms, window,start=50))) 
p$data=transform(merge(p$data,msy,by=c("qname","stock")),data=data/msy)
p$data=transform(p$data, quant=factor(quant,labels=c("Fishing \nMortality", "Yield", "Recruitment", "SSB")))

p$data$stock=factor(p$data$stock,
                    labels=c("sprat","brill","turbot","pollack","ray"),
                    levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"))

p=p+facet_grid(qname~stock,scale="free")+
  geom_hline(aes(yintercept=1),col="red",linetype=2)+
  theme_bw()+
  theme(legend.position="none")+
  xlab("Year")

p8=plot(FLStocks(llply(oms, function(x) window(iter(x,8),start=50))))
p8$data=transform(merge(p8$data,msy,by=c("qname","stock")),data=data/msy)
p8$data=transform(p8$data, quant=factor(quant,labels=c("Fishing \nMortality", "Yield", "Recruitment", "SSB")))
p8=p8+facet_grid(qname~stock,scale="free")+
  geom_hline(aes(yintercept=1),col="red",linetype=2)+
  theme_bw()+
  theme(legend.position="none")+
  xlab("Year")
p8$data$stock=factor(p8$data$stock,
                    labels=c("sprat","brill","turbot","pollack","ray"),
                    levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"))

p+geom_line(aes(year,data),data=p8$data,col="grey50")+
  facet_grid(quant~stock,scale="free")+
  scale_color_manual(values=rep("black",5))+
  scale_fill_manual(values=rep("grey",5))+
  theme_bw()+
  theme(legend.position="none", text=element_text(size=20))+
    facetted_pos_scales(
    y = list(
      quant == "Fishing \nMortality"~scale_y_continuous(limits=c(0, 2.5),  breaks=c(0,1.0,2.0)),
      quant == "Yield"              ~scale_y_continuous(limits=c(0, 2.0),  breaks=c(0,1.0)),
      quant == "Recruitment"        ~scale_y_continuous(limits=c(0, 3.0),  breaks=c(0,1.0,2.0)),
      quant == "SSB"                ~scale_y_continuous(limits=c(0, 12.5), breaks=c(0,0.1,10))))

```


\elandscape

```{r, lfd-data, eval=FALSE}
invAlks =llply(lhs,invAlk,cv=0.1)

lfds=mdply(data.frame(stock=as.character(names(oms)),stringsAsFactors=FALSE), 
           function(stock){
  om=window(oms[[stock]],start=2)
  
  if (any(any(catch.n(om)<0)))
    catch.n(om)[catch.n(om)<0]=0.0
  
  catch.n(om)[1]=0
  lfd=lenSample(catch.n(om),invAlks[[stock]],nsample=500)
  lfd=transform(melt(lfd),data=value)[,-4]
  lfd=subset(as.data.frame(lfd,drop=TRUE),data>0)
  
  lfd})

save(lfds,invAlk,file="data/lfds.RData")
```

\newpage
```{r, lfd, fig.width=12, fig.height=6}
load("/home/laurence-kell/Desktop/projects/mydas/papers/roc/data/lfds.RData")
load("/home/laurence-kell/Desktop/papers/roc/results/indicators.RData")

dat=ddply(subset(lfds,year%in%c(51,100,120)&iter==2),.(stock), transform,
          len =len/max(len),
          max =max(len))
dat$label=c("0.1Fmsy","2.5Fmsy","Fmsy")[factor(dat$year)]
names(dat)[1]="Stock"
max=dat[!duplicated(dat[,c("Stock")]),][,c("Stock","max")]

lns=subset(indicators,year%in%c(51,100,120))
lns=lns[!duplicated(lns[,c("Stock","year")]),
                   c("Stock","year","lmega","lopt","linf","l50","lc")]
                   
lns=merge(lns,max,by="Stock")
ln2=transform(lns,
          lmega =lmega/max,
          lopt  =lopt/max,
          linf  =linf/max,
          l50   =l50 /max,
          lc    =lc/max)
ln2=melt(ln2[,-(7:8)],id=c("Stock","year"))

dat$Stock=factor(dat$Stock,levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                           labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
ln2$Stock=factor(ln2$Stock,levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                           labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
dat$year=factor(dat$year,levels=c(51,100,120),
                         labels=c("FMSY","2.5FMSY","0.7FMSY"))
ln2$year=factor(ln2$year,levels=c(51,100,120),
                         labels=c("FMSY","2.5FMSY","0.7FMSY"))

ggplot(dat)+    
  geom_histogram(aes(len,weight=data),binwidth=0.05)+
  geom_vline(aes(xintercept=value,col=variable),data=ln2)+
  facet_grid(year~Stock,scale="free")+
  xlab("Relative Length")+ylab("")+
  scale_color_manual("Indicator",values=rainbow(5))+
  theme_bw()+
  theme(legend.position="none", text=element_text(size=20))
```

**Figure `r iFig=iFig+1; iFig`.** Simulated length frequencies distributions with indicators.

\blandscape
```{r, inds,fig.width=12,fig.height=8}
load("/home/laurence-kell/Desktop/papers/roc/results/indicators.RData")
indicators=subset(indicators,!(scen%in%dimnames(subset(design,sel2==5000))[[1]]))

quad=data.frame(x=c(50,80, 80, 50,  80,90,90,80, 90,105,105,90, 
                    105,110,110,105, 110,120,120,110), 
                y=rep(c(-Inf,-Inf,Inf,Inf),5),
                f=c(rep("green", 4),rep("amber", 4),rep("red",4),
                    rep("amber2",4),rep("green2",4)))
indicators=transform(subset(indicators,Fishery=="Independent"),
               l95   =(l95/linf),
               l25   =(l25/l50),
               lmax5 =(lmax5/linf),
               lmean =(lmean/lopt),
               lbar  =(lbar/l50),
               lmaxy =(lmaxy/lopt),
               lc    =(lc/l50),
               pmega =(pmega))

ref=data.frame(quant=c("l95","l25","lmax5","lmean","lbar","lmaxy","lc","pmega"),
               ref  =c(  0.8,  1.0,    0.8,    1.0,   1.0,    1.0, 1.0,   0.3))          
ref$quant=factor(ref$quant,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))

inds=subset(subset(indicators,scen%in%dimnames(subset(design,CV==0.3&sel2==1&sel3==5000&m=="gislason"&nsample==500&s==0.9))[[1]]&year>=50&year<=120))

lind=dlply(inds,.(Stock),with, FLQuants(
                  l25  =as.FLQuant(data.frame(year=year,iter=iter,data=l25),  units="NA"),
                  lc   =as.FLQuant(data.frame(year=year,iter=iter,data=lc),   units="NA"),
                  lmaxy=as.FLQuant(data.frame(year=year,iter=iter,data=lmaxy),units="NA"),
                  lmean=as.FLQuant(data.frame(year=year,iter=iter,data=lmean),units="NA"),
                  lbar =as.FLQuant(data.frame(year=year,iter=iter,data=lbar), units="NA"),
                  lmax5=as.FLQuant(data.frame(year=year,iter=iter,data=lmax5),units="NA"),
                  l95  =as.FLQuant(data.frame(year=year,iter=iter,data=l95) , units="NA"),
                  pmega=as.FLQuant(data.frame(year=year,iter=iter,data=pmega),units="NA")))
              
# Reduce FLQuants to FLQuant with names as quant
fqs <- FLQuants(lapply(lind, function(x) {
  res <- Reduce(f=qbind, x=x)
  dimnames(res)$quant <- names(x)
  return(res)
}))

# plot FLQuants w/ facet quant~qname
p=plot(fqs)
p$data$qname=factor(p$data$qname,levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                                 labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
p$data$quant=factor(p$data$quant,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))


p+facet_grid(quant~qname, scale="free")+
  geom_polygon(aes(x,y,fill=f), data=quad, alpha=0.2) +
  scale_fill_manual(values=c("orange","orange","orange","green","green","red"))+
  geom_vline(aes(xintercept=x),
    data=data.frame(y="2FMSY", x=c(90,105)), col="red") +
  geom_vline(aes(xintercept=x),
    data=data.frame(y="2FMSY", x=c(80,110)), col="orange") +
  geom_hline(aes(yintercept=ref),data=ref, linetype=2)+
  xlab("Year")+ylab("Indicator")  
```

**Figure `r iFig=iFig+1; iFig`.** Time series of fishery independent indicators relative to $F:F_{MSY}$, vertical lines indicate 1 (green), 1.5 (orange) and 2 (red) times $F_{MSY}$.


```{r, indc,fig.width=12,fig.height=8}
load("/home/laurence-kell/Desktop/papers/roc/results/indicators.RData")
indicators=subset(indicators,!(scen%in%dimnames(subset(design,sel2==5000))[[1]]))

indicators=transform(subset(indicators,Fishery=="Dependent"),
               l95   =(l95/linf),
               l25   =(l25/l50),
               lmax5 =(lmax5/linf),
               lmean =(lmean/lopt),
               lbar  =(lbar/l50),
               lmaxy =(lmaxy/lopt),
               lc    =(lc/l50),
               pmega =(pmega))

inds=subset(subset(indicators,scen%in%dimnames(subset(design,CV==0.3&sel2==1&sel3==5000&m=="gislason"&nsample==500&s==0.9))[[1]]&year>=50&year<=120))

lind=dlply(inds,.(Stock),with, FLQuants(
                  l25  =as.FLQuant(data.frame(year=year,iter=iter,data=l25),  units="NA"),
                  lc   =as.FLQuant(data.frame(year=year,iter=iter,data=lc),   units="NA"),
                  lmaxy=as.FLQuant(data.frame(year=year,iter=iter,data=lmaxy),units="NA"),
                  lmean=as.FLQuant(data.frame(year=year,iter=iter,data=lmean),units="NA"),
                  lbar =as.FLQuant(data.frame(year=year,iter=iter,data=lbar), units="NA"),
                  lmax5=as.FLQuant(data.frame(year=year,iter=iter,data=lmax5),units="NA"),
                  l95  =as.FLQuant(data.frame(year=year,iter=iter,data=l95) , units="NA"),
                  pmega=as.FLQuant(data.frame(year=year,iter=iter,data=pmega),units="NA")))

# Reduce FLQuants to FLQuant with names as quant
fqs <- FLQuants(lapply(lind, function(x) {
  res <- Reduce(f=qbind, x=x)
  dimnames(res)$quant <- names(x)
  return(res)
}))

# plot FLQuants w/ facet quant~qname
p=plot(fqs) +
  facet_grid(quant~qname, scale="free") +
  geom_polygon(aes(x,y,fill=f), data=quad, alpha=0.2) +
  geom_hline(aes(yintercept=ref),data=ref, linetype=2)+
  geom_vline(aes(xintercept=x),
    data=data.frame(y="2FMSY", x=c(90,105)), col="red") +
  geom_vline(aes(xintercept=x),
    data=data.frame(y="2FMSY", x=c(80,110)), col="orange") +
  scale_fill_manual(values=c("orange","orange","orange","green","green","red"))+
  xlab("Year")+ylab("Indicator")

head(p$data$quant)
p$data$qname=factor(p$data$qname,levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                                 labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
p$data$quant=factor(p$data$quant,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))

p
```

**Figure `r iFig=iFig+1; iFig`.** Time series of fishery dependent indicators relative to $F:F_{MSY}$, vertical lines indicate 1 (green), 1.5 (orange) and 2 (red) times $F_{MSY}$.


\elandscape

```{r}
load("/home/laurence-kell/Desktop/papers/roc/results/indicators.RData")
indicators=subset(indicators,!(scen%in%dimnames(subset(design,sel2==5000))[[1]]))  

ind=transform(subset(indicators,year%in%61:100),
              overfished=!(as.numeric(year)%in%80:100))
ind=ind[,c("Fishery","Stock","scen","overfished","l95","l25","lmax5","lmean","lbar","lmaxy","lc","pmega","lmega")]
ind=melt(ind,id=c("Fishery","Stock","scen","overfished"))

roc=ddply(ind,.(Stock,Fishery,scen,variable),with, 
                      mydas:::roc(overfished,value))

bst=ddply(roc,.(Stock,Fishery,scen,variable), transform, 
                     dst=(1-TPR)^2+(FPR)^2)
bst=ddply(bst,.(Stock,Fishery,scen,variable), with, 
                     data.frame(TPR      =TPR[dst==min(dst)],
                                FPR      =FPR[dst==min(dst)],
                                reference=reference[dst==min(dst)]))
names(bst)[4]="Indicator"

ices=transform(subset(indicators,year%in%61:100),
               overfished=(as.numeric(year)%in%80:100),
               l95   =(l95/linf-0.8)^2,
               l25   =(l25/l50-1)^2,
               lmax5 =(lmax5/linf-0.8)^2,
               lmean =(lmean/lopt-1)^2,
               lbar  =(lbar/l50-1)^2,
               lmaxy =(lmaxy/lopt-1)^2,
               lc    =(lc/l50-1)^2,
               pmega =(pmega-0.3)^2)
ices=melt(ices[,c("Fishery","Stock","scen","overfished","l95","l25","lmax5","lmean","lbar","lmaxy","lc","pmega","lmega")],
          id=c("Fishery","Stock","scen","overfished"))
ices=cbind(roc[ do.call("order",roc[,c("Fishery","Stock","scen","variable")]),],
           ices=ices[do.call("order",ices[,c("Fishery","Stock","scen","variable")]),"value"])

ices=ddply(ices,.(Stock,Fishery,scen,variable),with, data.frame(TPR  =TPR[ices==min(ices)][1],
                                                                FPR  =FPR[ices==min(ices)][1],
                                                                ices=ices[ices==min(ices)][1]))

roc$Stock =factor(roc$Stock, levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                             labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
bst$Stock =factor(bst$Stock, levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                             labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
ices$Stock=factor(ices$Stock,levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                             labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))

names(roc)[4] ="Indicator"
names(bst)[4] ="Indicator"
names(ices)[4]="Indicator"

roc =subset(roc, Indicator!="lmega")
bst =subset(bst, Indicator!="lmega")
ices=subset(ices,Indicator!="lmega")

roc$Indicator =factor(roc$Indicator, levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
bst$Indicator =factor(bst$Indicator, levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
ices$Indicator=factor(ices$Indicator,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
```


```{r, roc-overf}
ggplot(roc)+ 
  geom_path(aes(FPR,TPR,group=paste(scen,Fishery),col=Fishery))+
  facet_grid(Indicator~Stock)+
  geom_line(aes(x,y),data.frame(x=seq(0,1,0.01),y=seq(0,1,0.01)),linetype="dashed")+
  #geom_point(aes(FPR,TPR),data=bst,col="black")+
  #geom_point(aes(FPR,TPR),data=subset(bst,TPR+FPR>1),col="blue",size=2)+
  geom_point(aes(FPR,TPR,fill=Fishery),data=ices,shape=21,size=2.5)+
  xlab("False Positive Rate")+ylab("True Positive Rate")+
  scale_fill_manual(values=c("black","grey"))+
  scale_colour_manual(values=c("black","grey"))+
  theme_bw(16)+
  theme(legend.position="bottom") 

bst =transform(bst,score=((1-TPR)^2+FPR^2)^0.5)
bst =cbind(bst,design[bst$scen,])
ices=transform(ices,score=((1-TPR)^2+FPR^2)^0.5)
ices=cbind(ices,design[ices$scen,])     

save(roc,bst,ices,file="/home/laurence-kell/Desktop/papers/roc/results/overf.RData")  
```
**Figure `r iFig=iFig+1; iFig`.** ROC curves to detect overfishing i.e.  years 80 to 100 over 60 to 100


```{r}
load("/home/laurence-kell/Desktop/papers/roc/results/indicators.RData")
indicators=subset(indicators,!(scen%in%dimnames(subset(design,sel2==5000))[[1]]))

ind=transform(subset(indicators,year%in%100:120),
              overfished=!(as.numeric(year)%in%100:109))
ind=ind[,c("Fishery","Stock","scen","overfished","l95","l25","lmax5","lmean","lbar","lmaxy","lc","pmega")]
ind=melt(ind,id=c("Fishery","Stock","scen","overfished"))

roc=ddply(ind,.(Stock,Fishery,scen,variable),with, 
                      mydas:::roc(overfished,value))

bst=ddply(roc,.(Stock,Fishery,scen,variable), transform, 
                     dst=(1-TPR)^2+(FPR)^2)
bst=ddply(bst,.(Stock,Fishery,scen,variable), with, 
                     data.frame(TPR      =TPR[dst==min(dst)],
                                FPR      =FPR[dst==min(dst)],
                                reference=reference[dst==min(dst)]))
names(bst)[4]="Indicator"

ices=transform(subset(indicators,year%in%100:120),
               overfished=(as.numeric(year)%in%100:100),
               l95   =(l95/linf-0.8)^2,
               l25   =(l25/l50-1)^2,
               lmax5 =(lmax5/linf-0.8)^2,
               lmean =(lmean/lopt-1)^2,
               lbar  =(lbar/l50-1)^2,
               lmaxy =(lmaxy/lopt-1)^2,
               lc    =(lc/l50-1)^2,
               pmega =(pmega-0.3)^2)
ices=melt(ices[,c("Fishery","Stock","scen","overfished","l95","l25","lmax5","lmean","lbar","lmaxy","lc","pmega")],
          id=c("Fishery","Stock","scen","overfished"))
ices=cbind(roc[ do.call("order",roc[,c("Fishery","Stock","scen","variable")]),],
           ices=ices[do.call("order",ices[,c("Fishery","Stock","scen","variable")]),"value"])

ices=ddply(ices,.(Stock,Fishery,scen,variable),with, data.frame(TPR  =TPR[ices==min(ices)][1],
                                                                FPR  =FPR[ices==min(ices)][1],
                                                                ices=ices[ices==min(ices)][1]))
names(ices)[4]="Indicator"
names(roc)[ 4]="Indicator"

roc$Stock =factor(roc$Stock, levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                             labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
bst$Stock =factor(bst$Stock, levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                             labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))
ices$Stock=factor(ices$Stock,levels=c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata"),
                             labels=c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray"))

roc =subset(roc, Indicator!="lmega")
bst =subset(bst, Indicator!="lmega")
ices=subset(ices,Indicator!="lmega")

roc$Indicator=factor(roc$Indicator,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
bst$Indicator=factor(bst$Indicator,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
ices$Indicator=factor(ices$Indicator,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
```


```{r, roc-recover}
ggplot(roc)+ 
  geom_path(aes(FPR,TPR,group=paste(scen,Fishery),col=Fishery))+
  facet_grid(Indicator~Stock)+
  geom_line(aes(x,y),data.frame(x=seq(0,1,0.01),y=seq(0,1,0.01)),linetype="dashed")+
  #geom_point(aes(FPR,TPR),data=bst)+
  geom_point(aes(FPR,TPR,fill=Fishery),data=ices,shape=21,size=2.5)+
  scale_fill_manual(values=c("black","grey"))+
  scale_colour_manual(values=c("black","grey"))+
  xlab("False Positive Rate")+ylab("True Positive Rate")+
  theme(legend.position="none", text=element_text(size=20))+
  theme_bw(16)+
  theme(legend.position="bottom")

bst =transform(bst,score=((1-TPR)^2+FPR^2)^0.5)
bst =cbind(bst,design[bst$scen,])
ices=transform(ices,score=((1-TPR)^2+FPR^2)^0.5)
ices=cbind(ices,design[ices$scen,])

save(roc,bst,ices,file="/home/laurence-kell/Desktop/papers/roc/results/recover.RData")  
```

**Figure `r iFig=iFig+1; iFig`.** ROC recovery, power to detect reduction in fishing from 100 to 120


```{r, smry1-overf}
load("/home/laurence-kell/Desktop/papers/roc/results/design.RData")  
load("/home/laurence-kell/Desktop/papers/roc/results/overf.RData") 

names(ices)[4]="Indicator"
names(bst)[4] ="Indicator"
dat=rbind(cbind(Source="ICES", ices[,c("Stock","Fishery","scen","score","Indicator")]),
          cbind(Source="Tuned",bst[, c("Stock","Fishery","scen","score","Indicator")]))

design$Stock=factor(design$Stock,levels=rev(c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata")),
                                 labels=rev(c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray")))
dat=subset(dat,Indicator!="lmega")

dat$Indicator=factor(dat$Indicator,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))

dat=merge(dat,cbind(scen=seq(90),key=key,design))
dat$Stock=factor(dat$Stock,levels=rev(c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray")))
ggplot(dat)+
  geom_boxplot(aes(Stock,score,fill=Fishery))+
  scale_fill_manual(values=c("black","grey"))+
  theme(legend.position="right")+
  facet_grid(Indicator~Source)+
  coord_flip()+scale_y_continuous(limits=c(0,1.25))+
  ylab("Euclidean Distance Best Classification")+xlab("Indicator")+
  theme_bw(16)+
  theme(legend.position="bottom", text=element_text(size=20))
dat1=dat
```

**Figure `r iFig=iFig+1; iFig`.** Euclidean distance from TPR=1 and FPR=0 for detecting overfishing.

```{r smry1-recover}
load("/home/laurence-kell/Desktop/papers/roc/results/recover.RData")
load("/home/laurence-kell/Desktop/papers/roc/results/design.RData")

names(ices)[4]="Indicator" 
names(bst)[ 4]="Indicator"
dat=rbind(cbind(Source="ICES", ices[,c("Stock","Fishery","scen","score","Indicator")]),
          cbind(Source="Tuned",bst[, c("Stock","Fishery","scen","score","Indicator")]))

design$Stock=factor(design$Stock,levels=rev(c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata")),
                                 labels=rev(c("Sprat",                     "Brill",               "Turbot",       "Pollack",              "Ray")))
dat$Indicator=factor(dat$Indicator,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
dat=merge(dat,cbind(scen=seq(90),key=key,design))
dat$Stock=factor(dat$Stock,levels=rev(c("Sprat","Brill","Turbot","Pollack","Ray")))
ggplot(dat)+
  geom_boxplot(aes(Stock,score,fill=Fishery))+
  scale_fill_manual(values=c("black","grey"))+
  theme(legend.position="right")+
  facet_grid(Indicator~Source)+
  coord_flip()+scale_y_continuous(limits=c(0,1.25))+
  ylab("Euclidean Distance From Best Classification")+xlab("Indicator")+
  theme_bw(16)+
  theme(legend.position="bottom", text=element_text(size=20))
```

**Figure `r iFig=iFig+1; iFig`.** Euclidean distance from TPR=1 and FPR=0 for detecting recovery.


```{r smry1-recover-over}
ggplot(rbind(cbind(Stage="Overfish",dat1),cbind(Stage="Recovery",dat)))+
  geom_boxplot(aes(Stock,score,fill=Fishery))+
  scale_fill_manual(values=c("black","grey"))+
  theme(legend.position="right")+
  facet_grid(Indicator~Source+Stage)+
  coord_flip()+scale_y_continuous(limits=c(0,1.25))+
  ylab("Euclidean Distance From Best  Classification")+xlab("Indicator")+
  theme_bw(16)+
  theme(legend.position="bottom", text=element_text(size=20))
```

**Figure `r iFig=iFig+1; iFig`.** Euclidean distance from TPR=1 and FPR=0 


```{r, auc}
library(pROC)

load("/home/laurence-kell/Desktop/papers/roc/results/design.RData")
design$Stock=factor(design$Stock,levels=rev(c("Sprattus sprattus sprattus","Scophthalmus rhombus","Psetta maxima","Pollachius pollachius","Raja clavata")),
                                 labels=rev(c("Sprat",                     "Brill",              "Turbot",              "Pollack",              "Ray")))

load("/home/laurence-kell/Desktop/papers/roc/results/overf.RData")  
dat=cbind(Status="Over Fished",roc)
load("/home/laurence-kell/Desktop/papers/roc/results/recover.RData")
dat=rbind(dat,cbind(Status="Recovery",roc))
dat=subset(dat,Indicator!="lmega")
dat=ddply(dat,.(Stock,scen,Indicator,Fishery,Status), with, data.frame(auc=auc(labels, reference)))

dat$Indicator=factor(dat$Indicator,levels=c("lc","l25","lmaxy","lmean","lbar","lmax5","l95","pmega"))
dat=merge(dat,cbind(scen=seq(90),key=key,design))
dat$Stock=factor(dat$Stock,levels=rev(c("Sprat","Brill","Turbot","Pollack","Ray")))
ggplot(dat)+
  geom_boxplot(aes(Stock,auc,fill=Fishery))+
  geom_hline(aes(yintercept=c(0.5)))+
  geom_hline(aes(yintercept=c(0.7)))+
  geom_hline(aes(yintercept=c(0.8)))+
  geom_hline(aes(yintercept=c(0.9)))+
  scale_fill_manual(values=c("black","grey"))+
  theme(legend.position="right")+
  facet_grid(Indicator~Status)+
  coord_flip()+ 
  ylab("Area under the ROC curve")+xlab("Indicator")+
  theme_bw(16)+
  theme(legend.position="bottom", text=element_text(size=20))
```

**Figure `r iFig=iFig+1; iFig`.** Area under the ROC curve.

\newpage
```{r}
rsd=ddply(dat,.(Stock,Indicator,Fishery,Status), with, data.frame(scen=scen,residual=auc-mean(auc)))
olr=subset(rsd,residual<(-0.1))
subset(dat[dimnames(olr)[[1]],],Fishery=="Dependent")[,c(1,3,5,7)]
```


\newpage
```{r, smry2-overf}
load("/home/laurence-kell/Desktop/papers/roc/results/overf.RData")
names(bst)[4]="Indicator"  

tab=ddply(bst,.(Stock,Fishery,Indicator), with, data.frame(score=mean(score),TPR=mean(TPR),FPR=mean(FPR)))
tab=melt(tab[,c(1:3,5:6)],id=c("Stock","Fishery","Indicator"))
tab$Stock=factor(tab$Stock,levels=rev(c("Sprat","Brill","Turbot","Pollack","Ray")))

ggplot(tab)+
  geom_col(aes(x=value,y=Stock,fill=as.character(variable)))+
  scale_fill_manual(values=c("black","grey"))+
  facet_grid(Indicator~Fishery)+
  scale_fill_manual("Runs Test",values=c("red","black"),labels=c("FPR","TPR"))+
  xlab("Test")+
  scale_x_continuous(limits=c(0,1.5))+
  theme_bw(16)+
  theme(legend.position="bottom", text=element_text(size=20))
```
**Figure `r iFig=iFig+1; iFig`.** Overfishing, Summary of TPR & FPR

```{r, smry2-recover}
load("/home/laurence-kell/Desktop/papers/roc/results/recover.RData")
names(bst)[4]="Indicator"  

tab=ddply(bst,.(Stock,Fishery,Indicator), with, data.frame(score=mean(score),TPR=mean(TPR),FPR=mean(FPR)))
tab=melt(tab[,c(1:3,5:6)],id=c("Stock","Fishery","Indicator"))
     
tab$Stock=factor(tab$Stock,levels=rev(c("Sprat","Brill","Turbot","Pollack","Ray")))
ggplot(tab)+ 
  geom_col(aes(x=value,y=Stock,fill=as.character(variable)))+
  facet_grid(Indicator~Fishery)+
  scale_fill_manual("Runs Test",values=c("red","black"),labels=c("FPR","TPR"))+
  xlab("Test")+
  scale_x_continuous(limits=c(0,1.5))+
  theme_bw(16)+
  theme(legend.position="bottom", text=element_text(size=20))
```
**Figure `r iFig=iFig+1; iFig`.** Recovery, summary of TPR & FPR


```{r, tree-overf, fig.width=8,fig.height=8}
library(rpart)
library(rpart.plot)  

load("/home/laurence-kell/Desktop/papers/roc/results/overf.RData")

names(bst)[4]="Indicator"
dat=bst[,names(bst)[c(8,4,1:2,9:15)]]

fit=rpart(1/score~Indicator+Stock+Fishery+Stock+CV+s+sel2 + sel3 + nsample + m, data=dat, maxdepth=3)

rpart.plot(fit)
```
**Figure `r iFig=iFig+1; iFig`.** Overfishing, regression trees.


\newpage
```{r}
rpart.rules(fit,style="tall")  

class=cbind(class=fit[[2]],dat)
save(roc,bst,ices,class,file="/home/laurence-kell/Desktop/papers/roc/results/overf.RData")  
```  


\newpage
```{r, tree-recover, fig.width=8,fig.height=8}
library(rpart)  

load("/home/laurence-kell/Desktop/papers/roc/results/recover.RData")

names(bst)[4]="Indicator" 

dat=bst[,names(bst)[c(8,4,1:2,9:15)]]

fit=rpart(1/score~Indicator+Stock+Fishery+Stock+CV+s+sel2 + sel3 + nsample + m, data=dat, maxdepth=3)

rpart.plot(fit)
```

**Figure `r iFig=iFig+1; iFig`.** Recovery, regression trees.

\newpage
```{r}
rpart.rules(fit,style="tall")  
 
class=cbind(class=fit[[2]],dat) 
save(roc,bst,ices,class,file="/home/laurence-kell/Desktop/papers/roc/results/recover.RData")  
```


```{r, ref}
load("/home/laurence-kell/Desktop/papers/roc/results/indicators.RData")

ref=transform(subset(indicators,year==100),
               l95   =linf*0.8,
               l25   =l50,
               lmax5 =linf*0.8,
               lmean =lopt,
               lbar  =l50,
               lmaxy =lopt,
               lc    =l50,
               pmega =0.3)
ref=melt(ref[,c(1:2,9,12:20)],id=c("Stock","Fishery","scen"))
names(ref)[4]="Indicator"
names(nm)=spp
ref$Stock=nm[ref$Stock]

load("/home/laurence-kell/Desktop/papers/roc/results/overf.RData")

dat=subset(merge(bst[,c(1:4,7)],ref,by=c("Stock","Fishery","scen","Indicator")),!(Indicator%in%c("lmega","pmega")))

load("/home/laurence-kell/Desktop/papers/roc/results/recover.RData")
dt2=subset(merge(bst[,c(1:4,7)],ref,by=c("Stock","Fishery","scen","Indicator")),!(Indicator%in%c("pmega")))

dat=rbind(cbind(Stage="Overfished",dat),
          cbind(Stage="Recovery", dt2))
ggplot(transform(dat,ref=value/reference))+
  geom_boxplot(aes(Stock,ref,fill=Fishery))+
  theme(legend.position="right")+
  facet_grid(Indicator~Stage)+
  coord_flip()+ 
  ylab("Ratio")+xlab("")+
  theme_bw(16)+
  theme(legend.position="bottom", text=element_text(size=20))
```

\newpage 
# References {#References}

## Author information

**Laurence Kell**. laurie@seaplusplus.es

## Acknowledgements

This vignette and many of the methods documented in it were developed under the MyDas project funded by the Irish exchequer and EMFF 2014-2020. The overall aim of MyDas is to develop and test a range of assessment models and methods to establish Maximum Sustainable Yield (MSY) reference points (or proxy MSY reference points) across the spectrum of data-limited stocks.

## Software Versions

* `r version$version.string`
* FLCore:    `r packageVersion('FLCore')`
* FLBRP:     `r packageVersion('FLBRP')`
* FLasher:   `r packageVersion('FLasher')`
* FLife:     `r packageVersion('FLife')`
* mydas:     `r packageVersion('mydas')`
* **Compiled**: `r date()`


```{r}
flat=dimnames(subset(design,sel2==sel3))[[1]]

```

