---
title: "Performance Statistics"
subtitle: "Global MSY"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: ../refs.bib
editor_options: 
  markdown: 
    wrap: 72
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/html") })
---

When conducting Management Strategy Evaluation  objective need to be agreed and then quantified in the form of
statistics known as **Performance Statistics**.

An objective in many fisheries management frameworks is to achieve the maximum sustainable yield ($MSY$), therefore  the target biomass is $B_{MSY}$ and the target fishing mortality $F_{MSY}$. Where, $F_{MSY}$ is commonly defined as the fishing mortality with a given fishing pattern and current environmental conditions that gives the long-term maximum yield. 

If an objective is to achieve MSY, then $F_{MSY}$ and $B_{MSY}$ can be considered as target levels around which  fishing mortality and biomass should fluctuate around randomly, e.g. achieve the target levels with a 50% probability. 
However, the allocation of effort among fishing gears and between fleets is as important as the levl of fishing effort [@beddington1973356]. Differences in age-specific vulnerability to the fishing method can modify MSY, and so is rarely, if ever, attainable simultaneously for all species  [@maunder2002relationship]. An alternative is to use the
Global $MSY$, which is the yield that can be achieved for the optimal selection pattern. For example, by finding the knifed edged selction pattern that maximises yield.



```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path= "../cache/Brebuild/",
               fig.path  ="../output/tex/gmsy/",
               fig.width =8,
               fig.height=4,
               dev       ="png")
iFig=0
iTab=0
```

```{r, libs}
library( ggplot2); theme_set(theme_bw())

library(FLCore)
library(FLBRP)
library(ggplotFL)

require(FLasher)

library(plyr)
library(dplyr) 
library(reshape) 

library(interp)
```

```{r, eval=FALSE}
setwd("/home/laurie/Desktop/flr/FLCandy/examples/blim")
```

## Example using 'FLBRP' 

For a stock the 'FLBRP' object can be derived, this holds the expected values for the mortality, selectivity, mass, marurity-at-age, and the stock recruitment relationship, and can be used to calculate the expected, i.e. equilibrium values of biomass for any given level of fishing mortality. 


```{r, func}
setGeneric("globalMsy", function(object, ...)
  standardGeneric("globalMsy"))

# globalMsy {{{

#' @rdname globalMsy
#' @description Calculates the global MSY, i.e. the largest catch that can be made with a knife edge selection pattern
#' @examples
#' data(ple4brp)
#' globalMsy(ple4brp)
#' 

setMethod('globalMsy', signature(object='FLBRP'), function(object){
  globalMsyFn2(object)})

# }}}


globalMsyFn<-function(object){
    
  res=qapply(object,function(x) apply(x,c(1:5),median))
  res=propagate(res, dim(m(res))[1])
  
  range(res)[c("minfbar","maxfbar")]=range(res)["max"]

  mat=array(1,c(dim(m(object))[1],dim(m(object))[1]))
  mat[upper.tri(mat)]=0
  
  discards.sel(res)[]=0
  
  landings.sel(res)=FLQuant(c(mat),dimnames=dimnames(landings.sel(res)))
  
  refpts(res)=refpts(res)["msy",]
  refpts(res)[,"harvest"]=1
  msy=computeRefpts(res)["msy","yield",drop=T]
  msy[is.na(msy)]=0
  
  msy=seq(dim(m(res))[1])[msy==max(msy)][1]
  
  FLPar(c(age=an(dimnames(m(res))$age[msy]),computeRefpts(res)[,,msy,drop=T]))}


globalMsyFn2<-function(object){
  
  range(object)[c("minfbar","maxfbar")]=range(object)["max"]
  
  mat=array(1,c(dim(m(object))[1],dim(m(object))[1]))
  mat[upper.tri(mat)]   =0
  discards.sel(object)[]=0
  
  lsel=FLQuant(c(mat),dimnames=dimnames(propagate(iter(landings.sel(object),1),dim(m(object))[1])))
  
  refpts(object)=refpts(object)["msy",]
  
  rtn=NULL
  for (i in seq(dim(m(object))[1])){
    
    landings.sel(object)=iter(lsel,i)
    msy=computeRefpts(object)["msy","yield",drop=T]
    msy[is.na(msy)]=0
  
    msy=seq(dim(m(object))[1])[msy==max(msy)][1]
  
    res=cbind(age=an(dimnames(m(object))$age[i]),t(computeRefpts(object)[drop=T]))
    res=cbind(iter=an(dimnames(object)[[6]]),res)
    
    rtn=rbind(rtn,res)}

  rtn=as.data.frame(rtn)

  order=do.call("order",as.data.frame(rtn[,c("iter","harvest","age")]))
  rtn=rtn[rev(order),]
  rtn=rtn[!duplicated(rtn[,"iter"]),]

  order=do.call("order",as.data.frame(rtn[,"iter"]))
  rtn=rtn[order,]
  
  as(as.data.frame(rtn[,-1]),"FLPar")}
```


```{r, eq, fig.height=6,fig.width=8, eval=TRUE}
load("~/Desktop/inPrep/pew/github/erp/data/om/reference.RData")    

eq=brp(eq) 

gMSY=globalMsy(eq)["yield"]

gMSY

refpts(eq)["msy","yield"]
```






# References 
