---
title: "PSA priors"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR priors
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
# bibliography: refs.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir="html")})
---
  
```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)
opts_chunk$set(cache     =!TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="cache/era/",
               fig.path  ="../tex/era/",
               fig.width =8,
               fig.height=8,
               dev       ="png")
iFig=0
iTab=0
```


```{r, eval=FALSE, message=FALSE}
install.packages("devtools")

install.packages("ggplot2")
install.packages("ggpubr")
install.packages("GGallay")
install.packages("plyr")

devtools::install_github("flr/FLCore")
devtools::install_github("flr/ggplotFL")
devtools::install_github("flr/FLBRP")
devtools::install_github("flr/FLife")

devtools::install_github("henning-winker/FLRef")
devtools::install_github("henning-winker/SPMpriors")

devtools::install_github("ropensci/rfishbase")

# FLR
```

```{r, pkgs, cache=FALSE}
library(ggplot2)

library(FLCore)
library(ggplotFL)
library(FLBRP)
library(FLife)

library(rfishbase)

library(FishLife)
library(SPMpriors)
library(FLRef)

library(ggpubr)
library(GGally)

library(plyr)
library(reshape)

theme_set(theme_bw(16))
```


# Risk Equivalence 

Risk equivalence is an attempt to ensure that management advice and actions maintain a prescribed risk tolerance level across, stock and when uncertainty in the evaluation changes. Where risk is the probability that management targets will not be met or limits exceeded.


+ ICES Stock assessment categories	

  + **Category 1: stocks with quantitative assessments**  Full analytical assessments and forecasts as well as stocks with quantitative assessments based on production models.
  + **Category  2 stocks  with  analytical  assessments  that  are  only  treated  qualitatively**
  Quantitative assessments and forecasts which for a variety of reasons are considered indicative of trends in fishing mortality, recruitment, and biomass.
  + **Category 3: stocks for which survey based assessments indicate trends**
  Survey or other indices are available that provide reliable indications of trends in stock metrics, such as total mortality, recruitment, and biomass.
  + **Category 4: stocks for which only reliable catch data are available**
  Time series of catch can be used to approximate MSY.
  + **Category 5: landings only**
  Only landings data are available.
  + **Category 6: negligible landings**
  Landings are negligible in comparison to discards and stocks that are  primarily caught as bycatch species in other targeted fisheries 
  
+ Data quality categorisation (Patrick et al. 2009)
  
  + **Best data:** Information is based on collected data for the stock and area of interest that is established and substantial.
  Data rich stock assessment, published literature that uses multiple methods, etc.
  + **Adequate data:** Information with limited coverage and corroboration, or for some other reason deemed not as reliable as Tier 1 data
  Limited temporal or spatial data, relatively old information, etc
  + **Limited data:** Estimates with high variation and limited confidence and may be based on similar taxa or life history strategy.
  Similar genus or family, etc.
  + **Very limited data:** Expert opinion or based on general literature review from wide range of species, or outside of region
  General data – not referenced
  + **No data:** No information to base score on – not included in the PSA, but included in the DQI score.
  		 

## i) Get parameters from FishBase via FishLife
```{r, getTraits}
getTraits<-function(Class        ="predictive", 
                    Order        ="predictive", 
                    Family       ="predictive", 
                    Genus        ="predictive", 
                    Species      ="predictive",
                    partial_match=!TRUE,
                    add_ancestors=!TRUE, 
                    Database     =FishLife::FishBase_and_RAM, 
                    ParentChild_gz=Database$ParentChild_gz ){
  
  taxa=Search_species(Class,Order,Family,Genus,Species,add_ancestors,Database,ParentChild_gz)$match_taxonomy
  
  if(partial_match==TRUE)  Which=grep( taxa, ParentChild_gz[,'ChildName'])
  if(partial_match==FALSE) Which=which(taxa==ParentChild_gz[,'ChildName'])
  if( length(Which)!=1 ) stop( paste0("'Taxon' ",taxa," input matches more or less than one element") )

  Cov_gjj       =Database$Cov_gvv
  Mean_gj       =Database$beta_gv
  ParentChild_gz=Database$ParentChild_gz
  Y_ij          =Database$Y_ij
  g_i           =Database$g_i

  nms=c("linf","k","winf","amax","amat","m","l50","temperature","lnvar","rho",
        "lnmasps","lnmargsd","s","logits","fmsym","fmsy","lnr","r","lhg","g")
  mu =Mean_gj[Which,]
  names(mu)=nms
  mu =FLPar(mu)

  cov=Cov_gjj[Which,,]
  cov=as(array(cov,c(dim(cov),1),dimnames=list(params=nms,params=nms,iter=1)),"FLPar")
  
  FLPars(mu=mu,cov=cov)}
```


```{r, traits}
pars=getTraits(Genus="Solea",Species="solea")

par=c("linf","k","l50","s","m","lnmasps")
ref=c("s","m","r","fmsy","fmsym")

par(mfrow=c(2,2))
plot(  princomp(pars$cov[par,par,1,drop=T]),main=paste(par,collapse=","))
biplot(princomp(pars$cov[par,par,1,drop=T]))

plot(  princomp(pars$cov[ref,ref,1,drop=T]),main=paste(ref,collapse=","))
biplot(princomp(pars$cov[ref,ref,1,drop=T]))
```

**Figure `r iFig=iFig+1; iFig`** Summary of parameters


## ii) simulate multinormal distribution
```{r, mc}
set.seed(12234)

mu=pars$mu[ c("linf","k","l50","s"),1,drop=T]
cv=pars$cov[c("linf","k","l50","s"),c("linf","k","l50","s"),1,drop=T]
cvd=diag(cv)
cv[]=0
diag(cv)=cvd

par=mvtnorm::rmvnorm(100,mu,cv)
par=as(t(par),"FLPar")
dimnames(par)=list(params=names(mu),iter=seq(100))

par[c("linf","k","l50")]=exp(par[c("linf","k","l50")])

my_smooth <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_point(...,size=.5)+
  geom_smooth(...,method="lm",se=FALSE)}

my_density <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_density(...,lwd=1)}

ggpairs(model.frame(par)[,-5],
  lower = list(continuous = wrap(my_smooth)),
  diag=list(continuous=wrap(my_density,alpha=0.2)),
  title = "")+
  theme(legend.position ="none",
  panel.grid.major =element_blank(),
  axis.ticks       =element_blank(),
  axis.text.x      =element_blank(),
  axis.text.y      =element_blank(),
  panel.border     =element_rect(linetype = 1, colour="black", fill=NA))
```

**Figure `r iFig=iFig+1; iFig`** Pairwise scatter plots of life history parameters.

```{r}
data(teleost)

my_smooth <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_point(...,size=.5)+
  geom_smooth(...,method="lm",se=FALSE)}

my_density <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_density(...,lwd=1)}
  
ggpairs(cbind(transform(model.frame(teleost)[,-c(3,7)],
                        linf=log(linf),k=log(k),l50=log(l50))),
  lower = list(continuous = wrap(my_smooth)),
  diag=list(continuous=wrap(my_density,alpha=0.2)),
  title = "")+
  theme(legend.position ="bottom",
  panel.grid.major =element_blank(),
  axis.ticks       =element_blank(),
  axis.text.x      =element_text(angle=-45),
  axis.text.y      =element_blank(),
  panel.border     =element_rect(linetype = 1, colour="black", fill=NA))+
  theme_bw()
```


## iii) Fill missing with default parameters
```{r, lh}
lh=lhPar(par)
```

## iv) natural mortality
```{r, m, fig.height=4, fig.width=8}
dat=c(m1= 0.28, m1=(0.48- 0.07)/(2 * 1.96),
      m2=-1.30, m2=(- 1.19 +1.42)/(2 * 1.96),
      m3= 1.08, m3=(1.24 - 0.92)/(2 * 1.96))
mPar=FLPar(dat[seq(1,6,2)])
mSD =FLPar(dat[seq(2,6,2)])

set.seed(8899)
mPar=rbind(rnorm(100,mPar["m1"],mSD["m1"]),
           rnorm(100,mPar["m2"],mSD["m2"]),
           rnorm(100,mPar["m3"],mSD["m3"]))

lh=rbind(lh[!dimnames(lh)$params%in%c("m1","m2","m3")],mPar)

## v) Create FLBRP
mFn=function(object,par) {
   len=wt2len(stock.wt(object),par)

   #lnM= a + b ln(L/L∞) + c*lnK
   
   res= exp(par["m1"]%+%(par["m2"]%*%log(len/par["linf"]))%+%(par["m3"]%*%log(par["k"])))
   
   res}
```

## iv) Reference points and priors

```{r, eq}
eq =lhEql(lh, m=mFn)
```

```{r, fig.width=6,gig.height=10}
ggplot(FLQuants(mat=mat(eq), mass=stock.wt(eq), m=m(eq)), aes(x=age, y=data))+ 
  geom_flquantiles(fill="red",col="blue") + 
  facet_grid(qname~., scales='free')
```

```{r}
save(eq,file="~/Desktop/tmp/eq.RData")
```

**Figure `r iFig=iFig+1; iFig`** Biological parameters


```{r, ffmsy}
ggdensity(x="data", data=as.data.frame(refpts(eq)["msy","harvest"]))
```

**Figure `r iFig=iFig+1; iFig`** $F_{MSY}$.

```{r, eval=FALSE}
smeff=c("XIN","41","SQA","Cephalopods","Illex argentinus",        "Argentine shortfin squid",
        "CSW","34","OCC","Cephalopods","Octopus vulgaris",        "Common octopus",
        "CSW","34","DPS","Crustaceans","Parapenaeus longirostris","Deep-water rose shrimp",
        "CSW","47","ARV","Crustaceans","Aristeus varidens",       "Striped red shrimp",
        "CSW","47","GPW","Finfish",    "Epinephelus aeneus",      "White grouper",
        "CSW","34","GPW","Finfish",    "Epinephelus aeneus",      "White grouper",
        "XIN","41","HKX","Finfish",    "Merluccius spp",          "Hakes nei",
        "XIN","41","GRM","Finfish",    "Macruronus magellanicus", "Patagonian grenadier")
smeff=as.data.frame(t(array(smeff,c(6,8),list(variable=c("location", "FAO", "code","group","species","name"),.id=1:8))))
smeff=transform(smeff,genus  =substr(species,1,regexpr(" ", species)-1),
                      species=substr(species,  regexpr(" ", species)+1,nchar(species)))


siofa=c("Centroscymnus coelolepis","Pailona commun","Portugese dogfish","CYO",
        "Deania calcea","Squale savate","Birdbeak dogfish","DCA",
        "Centrophorus granulosus","Requin chagrin ","ulper shark","GUP",
        "Dalatias licha","Squale liche","Kitefin shark","SCK",
        "Bythaelurus bachi","Requin chat de Bach","Bach's catshark","BZO",
        "Chimaera buccanigella","Chimère bouche-foncée","Dark-mouth chimaera","ZZC",
        "Chimaera didierae","Chimère de Didier","The Falkor chimaera","ZZD",
        "Chimaera willwatchi","Chimère du marin","Seafarer's ghostshark","ZZE",
        "Centroscymnus crepidater","Pailona à long nez","Longnose Velvet Dogfish","CYP",
        "Centroscymnus plunketi","Pailona austral","Plunket shark","CYU",
        "Zameus squamulosus","Squale-grogneur à queue échancrée","Velvet dogfish","SSQ",
        "Etmopterus alphus","Requin lanterne à joues blanches","Whitecheek lanternshark","EZU",
        "Apristurus indicus","Holbiche artouca","Smallbelly catshark","APD",
        "Harriotta raleighana","Chimère à nez rigide","Bentnose rabbitfish","HCR",
        "Bythaelurus tenuicephalus","Requin chat à tête étroite","Narrowhead catshark","BZL",
        "Chlamydoselachus anguineus","Requin lézard","Frilled shark","HXC",
        "Hexanchus nakamurai","Requin griset","Bigeyed six-gill shark","HXN",
        "Etmopterus pusillus","Sagre nain","Smooth lanternshark","ETP",
        "Somniosus antarcticus","Requin dormeur antarctique","Southern sleeper shark","SON",
        "Mitsukurina owstoni","Requin lutin","Goblin shark","LMO")
siofa=as.data.frame(t(array(siofa,c(4,20))))
siofa=cbind(ldply(strsplit(siofa[,1]," ")),siofa)
names(siofa)=c("genus","species","scientificName","french","name","FAO")

dat=rbind.fill(smeff,siofa)

mdply(dat, function(genus,species) {
  res=try(FishLife::Search_species(Genus=genus,Species=species,add_ancestors=TRUE)$match_taxonomy)
  if ("try-error"%in%is(res)) {
    return(NULL)}
  res[1]}
  )

sn=with(subset(smeff,group!="Finfish"),paste(genus,species))

pg=popgrowth(    species_list=sn,server="sealifebase")[,c("Species","StockCode","Sex","Loo","K")]
mt=maturity(     species_list=sn,server="sealifebase")[,c("Species","StockCode","Sex","Lm")]
lw=length_weight(species_list=sn,server="sealifebase")[,c("Species","StockCode","Sex","a","b")]

pg=ddply(pg,.(Species), with, data.frame("loo"=mean(`Loo`,na.rm=T),"k"=mean(`K`,na.rm=T)))
mt=ddply(mt,.(Species), with, data.frame("l50"=mean(`Lm`,na.rm=T)))
lw=ddply(lw,.(Species), with, data.frame("a"=mean(`a`,na.rm=T),"b"=mean(`b`,na.rm=T)))

sl=merge(merge(pg,mt,by="Species"),lw,by="Species")

                                 
spp=data.frame(scientificName=c("Euthynnus affinis", "Thunnus tonggol", "Auxis rochei", "Auxis thazard"),
               code          =c("KAW",               "LOT",             "BLT",          "FRI"))
spp=transform(spp,genus  =substr(scientificName,1,regexpr(" ", scientificName)-1),
                  species=substr(scientificName,  regexpr(" ", scientificName)+1,nchar(scientificName)))

sardine
sprat
```

