---
title: "Introduction to FLCandy"
author: "Laurence Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to FLCandy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

FLCandy is an R package containing experimental methods and prototypes for the FLR (Fisheries Library for R) framework. This package provides implementations of various fisheries assessment and management tools that are under development or being evaluated for inclusion in the main FLR packages.

## Installation

```{r eval=FALSE}
# Install from GitHub
if (!require(devtools)) install.packages("devtools")
devtools::install_github("flr/FLCandy")
```

## Key Features

### Biological Reference Points

FLCandy provides several methods for calculating biological reference points:

```{r}
library(FLCandy)
library(FLCore)

# Load example data
data(ple4)
data(ple4brp)

# Calculate production type parameters
pt_params <- pt(ple4brp)
print(pt_params)
```

### Biological Indicators

Calculate a suite of biological indicators for stock assessment:

```{r}
# Calculate indicators
indicators <- ind(ple4, ple4brp)
head(indicators)
```

### Harvest Control Rules

Apply ICES-style harvest control rules:

```{r eval=FALSE}
# Apply harvest control rule
tac <- hcrICES(ple4, ple4brp, 
               start = 2010, 
               end = 2020)
```

## Utility Functions

FLCandy includes utility functions for parameter transformations:

```{r}
# Convert steepness to logit scale for optimization
h_values <- c(0.3, 0.5, 0.8)
logit_h <- to_logits(h_values)
print(logit_h)

# Convert back to original scale
original_h <- from_logits(logit_h)
print(original_h)
```

## Spatial Analysis

The package includes functions for spatial analysis of fisheries data:

```{r eval=FALSE}
# Spatial functions are available for geographic analysis
# See ?spatial_funs for more details
```

## Contributing

FLCandy contains experimental methods. If you find bugs or have suggestions for improvements, please:

1. Check existing issues on GitHub
2. Create a new issue with a reproducible example
3. Consider contributing code improvements

## References

For more information about the FLR framework, see:

- FLR website: https://www.flr-project.org/
- FLR documentation: https://flr-project.org/doc/ 