---
title: "**Predation Mortality**"
subtitle: "M1 & M2"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
---


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =!FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/m12/",
               fig.path  ="../outputs/tex/m12",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)    # install.packages("FLCore",   repos="http://flr-project.org/R")
library(FLBRP)     # install.packages("FLBRP",    repos="http://flr-project.org/R")
library(FLasher)   # install.packages("FLasher",  repos="http://flr-project.org/R")
library(FLife)     # install.packages("FLife",    repos="http://flr-project.org/R")
library(ggplotFL)  # install.packages("ggplotFL", repos="http://flr-project.org/R")

theme_set(theme_bw(16))
```

```{r, methods}
setGeneric("m1", function(object, ...)
  standardGeneric("m1"))
setGeneric("m2", function(object, ...)
  standardGeneric("m2"))
setGeneric("forage", function(object, ...)
  standardGeneric("forage"))
setGeneric("predNeed", function(object, ...)
  standardGeneric("predNeed"))

setMethod("z", "FLBRP", function(object, ...) {
  f <- harvest(object)
  if(units(f) != 'f')
    stop("Your exploitation rate is not defined as F, cannot be added to M")
  else 
    return(m(object) %+% f)
  })

setMethod("m1", "FLComp", function(object, ...) {
  if(units(harvest(object)) != 'f') 
    stop("Exploitation not defined as 'f', cannot be combined with  'm'")
  if (!"m1"%in%names(attributes(stk)))
    stop("m1 is not an attribute")
 
  return(attributes(object)$m1)})

setMethod("m2", "FLComp", function(object, ...) {
  if(units(harvest(object)) != 'f') 
    stop("Exploitation not defined as 'f', cannot be combined with  'm'")
  if (!"m1"%in%names(attributes(stk)))
    stop("m1 is not an attribute")
  
  return(m(object) - m1(object))})

setMethod("forage", "FLComp", function(object, ...) {
  if(units(harvest(object)) != 'f') 
    stop("Exploitation not defined as 'f', cannot be combined with  'm'")
  if (!"m1"%in%names(attributes(stk)))
    stop("m1 is not an attribute")

  #zFn<-function(object) m(object)%+%harvest(object)
  
  return(FLCore:::apply((stock.wt(object)%*%stock.n(object)%*%(m2(object))%/%(z(object))%*%(1-exp(-z(object))))[-1],c(2,6),sum))})

setMethod("predNeed", "FLBRP", function(object, ...) {
  
  # reference points
  rfs=refpts(object)[,"harvest"]
  
  # F reference points
  hvt=aperm(rfs,c(2,1,3:length(dim(rfs))))
  
  # Change to FLQuant  
  dimnames(hvt)[1:2]=list(age="all","year"=seq(dim(hvt)[2]))
  names(dimnames(hvt))[2]="year"
  fbar(object)=FLQuant(hvt)
    
  # Calc forage
  dimnames(rfs)[[2]]="forage"
  rfs[]=c(forage(object))
    
  rfs})
```


```{r, lh}
lh=FLPar(linf=25)
lh=lhPar(lh)
```

```{r, eq}
eq=lhEql(lh)

plot(eq,refpts="msy")
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium dynamics 

```{r, stk}
stk=as(eq,"FLStock")
stk=qapply(stk, function(object) {dimnames(object)$year=an(dimnames(object)$year)+1960; object})
stk=fwd(stk,f=fbar(stk)[,-1],sr=eq)

attributes(eq)$m1 =FLQuant(0.025,dimnames=dimnames(m(eq)))
attributes(stk)$m1=FLQuant(0.025,dimnames=dimnames(m(stk)))
```

```{r, ts}
ggplot(model.frame(FLQuants(eq,ssb=ssb,yield=catch,forage=forage),drop=T))+
  geom_line(aes(ssb,forage),col="red")+
  geom_line(aes(ssb,yield),col="blue")+
  xlab("SSB")+ylab("Production")
```

**Figure `r iFig=iFig+1; iFig`** Yiele verses Forage

```{r, yield-forgage}
plot(stk, metrics=list(F=fbar,SSB=ssb,Yield=catch,Forage=forage))+
  geom_vline(aes(xintercept=1995),col="red")+
  xlab("Year")
```

**Figure `r iFig=iFig+1; iFig`** Simulated time series.

```{r, need}
predNeed(eq)
```

## Software Versions

* `r version$version.string`
* FLCore:    `r packageVersion('FLCore')`
* FLBRP:     `r packageVersion('FLCore')`
* FLasher:   `r packageVersion('FLCore')`
* FLife:   `r packageVersion('FLCore')`
* ggplotFL:  `r packageVersion('FLCore')`

* FLPKG: `r # packageVersion('FLPKG')`

* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence KELL**. laurie.kell.es


## Acknowledgements


# References {#References}
