---
title: "Generation Time in Fish Populations"
subtitle: ""
author: " L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
  number_sections: false
toc: false
keep_tex: false
mathjax: TRUE
vignette: >
  %\VignetteIndexEntry{JABBA Guidelines}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE,
               eval      =FALSE,
               fig.width =10, 
               fig.height=4,
               cache     =TRUE, 
               fig.path  ="../tex/gt-",
               cache.path="cache/gt/",
               dev       ="png")

iFig=0
iTab=0
```


# Generation time 

Generation time is useful for comparing population dynamics across life-history 
types, conducting phylogenetic analyses, and estimating extinction risk. 

There are several ways to calculate generation time depending on the available data,
for example using life history parameters or age-based models. In the later case 
generation time can be estimated as the average age of parents of the recruits in 
an unfished population, in the former generation time can be estimated as the age 
at an optimal length ($L_{opt}$). For example the length at which a cohort reaches 
it's maximum exploitable or spawning biomass, when fish are most valuable for commercial 
purposes, or maximise their reproductive potential.

## Life-history

This can be calculated using growth models, such as the von Bertalanffy Growth Function (VBGF @vonB)

The **von Bertalanffy Growth Function (VBGF)** is a widely used equation in fisheries science to model the growth of fish and other animals. The equation describes how the length of an organism changes with age.

### Von Bertalanffy Growth Function (VBGF) Equation:

$L(t) = L_\infty \left(1 - e^{-k(t - t_0)}\right)$

Where, $L(t)$ is the length at age $t$, $L_\infty$ is the asymptotic maximum length 
(the maximum theoretical length the organism can reach), $k$ is the growth rate 
coefficient (how quickly the organism approaches $L_\infty$), and $t_0$ is the 
theoretical age at which the organism would have zero length.

If the optimal length is knowm, you can rearrange the equation to solve for age $t$:

$t = t_0 - \frac{1}{k} \ln \left(1 - \frac{L(t)}{L_\infty}\right)$


```{r, age}
library(FLife)

par=FLPar(linf=80,k=0.2,t0=0.2)

vonB(age=5,par)
```

$L_{opt}$ is a function of growth and natural mortality-at-age and there are several 
 approximations such as $L_{\infty}\frac{3}{3+k/m}$ and 2/3$L_{\infty}$



## Leslie Matrix Framework


A Leslie matrix is a discrete, age-structured model that describes population growth where:

```R
# Basic Leslie Matrix Structure
L <- matrix(0, nrow = n, ncol = n)
# First row contains fecundity rates (Fx)
L[1, ] <- c(F1, F2, ..., Fn)
# Subdiagonal contains survival rates (Sx)
for(i in 1:(n-1)) {
  L[i+1, i] <- Sx[i]
}
```

## Generation Time Calculation Methods

There are three main approaches to estimate generation time from a Leslie matrix:

1. **Method A (Age-weighted)**
$ A = \sum_{j=1}^n jp_j\lambda^{-j} $
Where:
- j is the age class
- pj is the reproductive value
- $\lambda$ is the population growth rate

2. **Method T (Time-weighted)**
$ T = \frac{\ln(R_0)}{\ln(\lambda)} $
Where:
- R0 is the net reproductive rate
- $\lambda$ is the dominant eigenvalue of the Leslie matrix

3. **Method m (Mean parent age)**
$ m = \sum_{j=1}^n jp_j $


```{r}
# Example data for a fish population
survival =c(0.5, 0.4,   0.3, 0.2, 0.2, 0.2)
fecundity=c(0,   0,   100, 200, 200, 200, 200)

leslieMatrix<-function(survival, fecundity) {
  n= length(survival)+1
  L= matrix(0, nrow=n, ncol=n)
  
  L[1,]=fecundity
  
  for(i in 1:(n-1))
    L[i+1, i] = survival[i]
  
  return(L)}

R0<-function(L) {
  # Fertility rates
  fertility=L[1, ]
  
  # Cumulative survival probabilities
  survival=cumprod(c(1, diag(L[-1, -ncol(L)])))
  
  # Calculate R0 as sum of products of survival and fertility
  R0 = sum(fertility*survival)
  
  return(R0)}

# Method A: Age-weighted generation time
gtA<-function(L) {
  
  lambda=Re(eigen(L)$values[1])
  w     =Re(eigen(L)$vectors[,1])
  v     =Re(eigen(t(L))$vectors[,1])
  
  # Normalize reproductive value vector
  v     =v/sum(v*w)
  
  ages=1:nrow(L)
  
  A=sum(ages*v*lambda^(-ages))
  
  return(A)}

# Method T: Time-weighted generation time
gtT<-function(L) {
  lambda=Re(eigen(L)$values[1])
  r0    =R0(L)
  T     =log(r0)/log(lambda)
  
  return(T)}

# Method m: Mean parent age
gtM<-function(L) {
  ages = 1:nrow(L)
  v = Re(eigen(t(L))$vectors[,1])
  m = sum(ages*v)/sum(v)
  
  return(m)}


# Create Leslie matrix
L=leslieMatrix(survival, fecundity)

# Calculate generation time using all methods
A=gtA(L)
T=gtT(L)
M=gtM(L)

# Compare results
results=data.frame(
  Method          =c("Age-weighted (A)", "Time-weighted (T)", "Mean parent age (m)"),
  Generation_Time =c(A,                   T,                   M))

results
```
