---
title: "Stock Analysis Functions in FLCandy"
author: "FLCandy Package"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Stock Analysis Functions in FLCandy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

# Introduction

The `FLCandy` package provides a comprehensive set of functions for analyzing individual stock trends and states using S4 methods. These functions can work with various input types including `FLStock` objects, `FLQuant` objects, and data frames.

## Key Features

- **S4 Method Dispatch**: Functions automatically adapt to different input types
- **Multiple Analysis Methods**: Linear regression and Mann-Kendall trend tests
- **Year-by-Year Classification**: Sliding window approach for temporal analysis
- **Flexible Input**: Support for different data structures
- **Comprehensive Output**: Detailed trend information, p-values, and classifications

## Functions Overview

1. **`stockTrend`**: Analyze trends in SSB and F for a single stock
2. **`stockState`**: Classify stock state (Increasing, Decreasing, Stable) for a single stock
3. **`stockStateFactor`**: Create ordered factors for statistical analysis

# Loading Required Packages

```{r load-packages}
library(FLCandy)
library(FLCore)
library(ggplot2)

# Load example data
data(ple4)
```

# Single Stock Analysis

## Using FLStock Objects

The most straightforward way to analyze a stock is using an `FLStock` object:

```{r flstock-analysis}
# Basic trend analysis
trend_result <- stockTrend(ple4)
print(trend_result$trend)
print(trend_result$classification)

# Get stock state
state <- stockState(ple4)
print(state)

# Create factor for statistical analysis
state_factor <- stockStateFactor(ple4)
print(state_factor)
```

## Using FLQuant Objects

You can also analyze stocks using separate `FLQuant` objects for SSB and F:

```{r flquant-analysis}
# Extract SSB and F from FLStock
ssb_data <- ssb(ple4)
f_data <- fbar(ple4)

# Analyze using FLQuant objects
trend_result <- stockTrend(ssb = ssb_data, f = f_data)
print(trend_result$trend)

state <- stockState(ssb = ssb_data, f = f_data)
print(state)
```

## Using Data Frames

The functions also work with data frames containing the required columns:

```{r dataframe-analysis}
# Create data frame from FLStock
df_data <- data.frame(
  year = as.numeric(dimnames(ssb(ple4))$year),
  ssb = c(ssb(ple4)),
  f = c(fbar(ple4))
)

# Analyze using data frame
trend_result <- stockTrend(df_data)
print(trend_result$trend)

state <- stockState(df_data)
print(state)
```

# Year-by-Year Analysis

The functions support year-by-year classification using a sliding window approach:

```{r yearly-analysis}
# Year-by-year trend analysis
yearly_trend <- stockTrend(ple4, byYear = TRUE)
print(head(yearly_trend$yearlyStates))

# Year-by-year state classification
yearly_state <- stockState(ple4, byYear = TRUE)
print(head(yearly_state))
```

## Visualizing Year-by-Year Results

```{r yearly-plot, fig.width=10, fig.height=6}
# Plot yearly states over time
yearly_data <- stockState(ple4, byYear = TRUE)
yearly_df <- yearly_data[, c("year", "state")]

ggplot(yearly_df, aes(x = year, fill = state)) +
  geom_bar() +
  scale_fill_manual(values = c("Decreasing" = "red", "Stable" = "orange", "Increasing" = "green")) +
  labs(title = "Stock State Over Time", x = "Year", y = "Count", fill = "State") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Advanced Analysis Options

## Different Analysis Methods

```{r methods}
# Linear regression (default)
linear_result <- stockTrend(ple4, method = "linear")
print(linear_result$trend)

# Mann-Kendall test
mk_result <- stockTrend(ple4, method = "mann_kendall")
print(mk_result$trend)
```

## Custom Parameters

```{r parameters}
# Custom significance threshold
strict_result <- stockTrend(ple4, pThreshold = 0.01)
print(strict_result$trend)

# Custom window size for year-by-year analysis
custom_yearly <- stockTrend(ple4, byYear = TRUE, windowSize = 10)
print(head(custom_yearly$yearlyStates))

# Focus on recent years
recent_result <- stockTrend(ple4, recentYears = 20)
print(recent_result$trend)
```

## Detailed Results

```{r detailed-results}
# Get detailed results instead of simplified output
detailed <- stockState(ple4, simplify = FALSE)
print(names(detailed))
print(detailed$ssbTrend)
print(detailed$fTrend)
```

# Method Dispatch Examples

## Automatic Method Selection

The S4 methods automatically dispatch based on the input type:

```{r method-dispatch}
# FLStock method
result1 <- stockTrend(ple4)
print(class(result1))

# FLQuant method
result2 <- stockTrend(ssb = ssb(ple4), f = fbar(ple4))
print(class(result2))

# Data frame method
df <- data.frame(year = 1:10, ssb = rnorm(10), f = rnorm(10))
result3 <- stockTrend(df)
print(class(result3))
```

# Error Handling

The functions provide informative error messages for unsupported input combinations:

```{r error-handling, error=TRUE}
# This will generate an error - FLQuant needs both ssb and f
tryCatch({
  stockTrend(ssb = ssb(ple4))
}, error = function(e) {
  cat("Error:", e$message, "\n")
})

# This will generate an error - data frame needs required columns
tryCatch({
  stockTrend(data.frame(year = 1:10, wrong_column = rnorm(10)))
}, error = function(e) {
  cat("Error:", e$message, "\n")
})
```

# Performance Considerations

## Large Datasets

For large datasets, consider using the `recentYears` parameter to focus on recent data:

```{r performance}
# Analyze only recent 20 years for better performance
recent_trend <- stockTrend(ple4, recentYears = 20)
print(recent_trend$trend)
```

## Understanding the Output

The functions return detailed information about stock trends:

```{r output-explanation}
# Get comprehensive trend analysis
result <- stockTrend(ple4)

# Overall trend classification
cat("Overall trend:", result$trend, "\n")

# SSB trend details
cat("SSB trend direction:", result$ssbTrend$direction, "\n")
cat("SSB p-value:", result$ssbTrend$p_value, "\n")
cat("SSB significant:", result$ssbTrend$significant, "\n")

# F trend details
cat("F trend direction:", result$fTrend$direction, "\n")
cat("F p-value:", result$fTrend$p_value, "\n")
cat("F significant:", result$fTrend$significant, "\n")

# Classification explanation
cat("Classification:", result$classification, "\n")
```

# Summary

The stock analysis functions in `FLCandy` provide a comprehensive and flexible toolkit for analyzing individual stock trends and states. Key features include:

- **S4 Method Dispatch**: Automatic adaptation to different input types (FLStock, FLQuant, data.frame)
- **Multiple Analysis Methods**: Linear regression and Mann-Kendall tests
- **Temporal Analysis**: Year-by-year classification with sliding windows
- **Flexible Input**: Support for different data structures
- **Statistical Integration**: Factor creation for statistical analysis
- **Comprehensive Output**: Detailed trend information with p-values and classifications

These functions make it easy to analyze stock trends and states for individual stocks across different data formats, providing both simple classifications and detailed statistical information for further analysis. 