---
title: "Performance Metric"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE}
library(knitr)
source("R/ini.R")
```


## Required packages

To follow this tutorial you should have installed the following packages:

- CRAN: 
- FLR: [FLCore](http://www.flr-project.org/FLCore/)

You can do so as follows:

```{r, eval=FALSE}
install.packages(c("iterators"))
install.packages(c("FLCore"), repos="http://flr-project.org/R")
```


```{r, eval=FALSE}
library(devtools)

install_github("lauriekell/FLCandy")
```

# Section

```{r}
library(plyr)
library(reshape)
library(randtests)

```

```{r, flcore}
library(FLCore)
library(FLBRP)
library(FLasher)
library(ggplotFL)
```

```{r, FLCandy}
library(FLCandy)
library(mydas) 
library(FLife) 
```

```{r, data}
data("ple4")
data("ple4brp")
```


```{r, projections}
om=fwdWindow(ple4,end=2050,ple4brp)
deviates=rlnoise(100,rec(om)%=%1,0.3,0.7)
om=propagate(om,100)

om.msy=fwd(om,fbar=fbar(om)[,ac(2018:2050)]%=%c(refpts(ple4brp)["msy","harvest"]),sr=ple4brp,residuals=deviates)

om.f01=fwd(om,fbar=fbar(om)[,ac(2018:2050)]%=%c(refpts(ple4brp)["f0.1","harvest"]),sr=ple4brp,residuals=deviates)

om.mey=fwd(om,fbar=fbar(om)[,ac(2018:2050)]%=%c(refpts(ple4brp)["mey","harvest"]),sr=ple4brp,residuals=deviates)
```


```{r, projections-plots}
oms=FLStocks("MSY" =om.msy,
              "F0.1"=om.f01,
              "MEY" =om.mey)

plot(oms)
```


```{r, stats}
stats=ldply(oms,function(x) omSmry(x,ple4brp))
head(stats)
```


```{r, metrics}
pMetrics=ddply(stats, .(.id), mydas:::smryStat)
pMetrics
```


## Subsection


### More information

To learn more about this subsection, check the [FLPKG](http://www.flr-project.org/FLPKG/).

# References

@kell2017flr

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Alternatively, send a pull request to <https://github.com/flr/doc/>.
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage: <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author(s)

**Name SURNAME**. Institution, address, 00000 Place, Country <https://website.dom>.
