---
title: "Effort Trends"
subtitle: "Operating Model Conditioning"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
editor_options: 
  markdown: 
    wrap: 72
bibliography: ../tex/refs.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs/html") })
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path= "cache/OMeffort/",
               fig.path  ="../tex/figs/OMeffort/",
               fig.width =8,
               fig.height=4,
               dev       ="png")
iFig=0
iTab=0
```

```{r, libs}
library(ggplot2); theme_set(theme_bw())

library(FLCore)
library(FLBRP)
library(ggplotFL)

library(FLasher)

library(FLCandy)

library(plyr)
```

```{r, eval=FALSE}
setwd("~/Desktop/flr/FLCandy/examples/effort/Rmd")
```

```{r}
load("~/Desktop/inPrep/biomassrebuild/data/Updated_stks_n81.rdata")
```

## ICES Effort trends

@bell2017global

```{r}
icesF=ldply(stks,function(x) as.data.frame(fbar(x),drop=T))

ggplot(icesF)+
  geom_line(aes(year,data,group=.id))
```

## Global Effort trends
```{r}
year=seq(1950,2012)

asia=data.frame(year=c(1950,1970,1995,2012),
                E   =c(10,    10, 200, 250)/250)
asia=merge(cbind(asia,.id=seq(4)+1),cbind(asia,.id=seq(4)),by=".id")
asia=ddply(asia, .(.id), with, data.frame(F=seq(E.x,E.y,length.out=year.y-year.x+1),year=seq(year.x,t=year.y)))

europe=data.frame(year=c(1950,1960,1980,1990,2012),
                  E     =c(  10,  10,  15,  15, 10)/15)
europe=merge(cbind(europe,.id=seq(5)+1),cbind(europe,.id=seq(5)),by=".id")
europe=ddply(europe, .(.id), with, data.frame(F=seq(E.x,E.y,length.out=year.y-year.x+1),year=seq(year.x,t=year.y)))

africa=data.frame(year=c(1950,1990,2012),
                  E     =c(5,    5,    40)/40)
africa=merge(cbind(africa,.id=seq(3)+1),cbind(africa,.id=seq(3)),by=".id")
africa=ddply(africa, .(.id), with, data.frame(F=seq(E.x,E.y,length.out=year.y-year.x+1),year=seq(year.x,t=year.y)))

samerica=data.frame(year=c(1950,1980, 1990,  2012),
                     E       =c(   1,   1,  1.5,   7.5)/7.5)
samerica=merge(cbind(samerica,.id=seq(4)+1),cbind(samerica,.id=seq(4)),by=".id")
samerica=ddply(samerica, .(.id), with, data.frame(F=seq(E.x,E.y,length.out=year.y-year.x+1),year=seq(year.x,t=year.y)))

namerica=data.frame(year=c(1950,1970,  1975,1980,2012),
                    E   =c(   5,   5,   2.5, 2.5,  10)/10)
namerica=merge(cbind(namerica,.id=seq(5)+1),cbind(namerica,.id=seq(5)),by=".id")
namerica=ddply(namerica, .(.id), with, data.frame(F=seq(E.x,E.y,length.out=year.y-year.x+1),year=seq(year.x,t=year.y)))

oceania=data.frame(year=c(1950,1970,2000,2012),
                   E       =c( 2.5, 1, 1.25, 3)/3)
oceania=merge(cbind(oceania,.id=seq(4)+1),cbind(oceania,.id=seq(4)),by=".id")
oceania=ddply(oceania, .(.id), with, data.frame(F=seq(E.x,E.y,length.out=year.y-year.x+1),year=seq(year.x,t=year.y)))

F=rbind(cbind("Region"="Asia",         asia),
        cbind("Region"="Oceania",      oceania),
        cbind("Region"="South America",samerica),
        cbind("Region"="North America",namerica),
        cbind("Region"="Europe",       europe),
        cbind("Region"="Africa",       africa)
        )[,-2]

ref =ddply(subset(F,year>=1980), .(Region), with,  data.frame(year=max(year[(F-0.5)^2==min((F-0.5)^2)])))
ref2=ddply(subset(F,year>=1980), .(Region), with,  data.frame(year=max(year[(F-0.6)^2==min((F-0.6)^2)])))
```



The exploitation rate (E) is used as indicator expressing the status of stocks, 
a stock is considered as under-exploited if E<0.5, over-exploited if E>0.6 and
the target E is in the range 0.5-0.6.

```{r}
ggplot(subset(F,Region!="Europe"))+
  geom_polygon(aes(x,y,fill=Zone),data=data.frame(x   =rep(c(1950,2012,2012,1950,1950),3),
                                                  y   =c(0,0,1,1,0, 1,1,1.2,1.2,1, 1.2,1.2,2,2,1.2),
                                            Zone=rep(c("UnderFished","Sustainable","Overfished"),each=5)))+
  geom_line( aes(year,F*2,col=Region),size=2)+
  geom_point(aes(year,F*2),data=merge(F,ref))+
  scale_fill_manual(values=c("red","green","blue"))+
  xlab("Year")+ylab(expression(F:F[MSY]))
```

## Synthetic

```{r}
f=FLQuant(c(rep(0.1,20),seq(0.1,2,length.out=40),seq(2,0.7,length.out=5)[-1],rep(0.7,30)))

dimnames(f)=list(year=an(dimnames(f)$year)+1950)

plot(f)
```

## Software Versions

-   `r version$version.string`
-   FLCore: `r packageVersion('FLCore')`
-   FLPKG: `r # packageVersion('FLPKG')`

-   FLBRP:    `r packageVersion('FLBRP')`
-   FLasher:  `r packageVersion('FLasher')`
-   FLCandy:  `r packageVersion('FLCandy')`
-   ggplotFL: `r packageVersion('ggplotFL')`

-   **Compiled**: `r date()`
-   **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence KELL**. [laurie\@seaplusplus.co.uk](mailto:laurie@seaplusplus.co.uk){.email}

# References

