---
title: "FLife"
subtitle: "Operating Model Conditioning"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   pdf_document:
     keep_tex: true
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
editor_options: 
  markdown: 
    wrap: 72
bibliography:  refs.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs/html") })
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path= "cache/FLifeOM/",
               fig.path  ="../tex/figs/FLifeOM/",
               fig.width =8,
               fig.height=4,
               dev       ="png")
iFig=0
iTab=0
```

```{r, libs}
library(ggplot2); theme_set(theme_bw())

library(FLCore)
library(FLBRP)
library(ggplotFL)

library(FLasher)

library(FLife)
library(mydas)
library(FLCandy)

library(plyr)
```

When conducting Management Strategy Evaluation the Operating Model (OM)
describes the resource dynamics for use in simulation to provide a basis
for the comparison of alternative assessment methods and management
strategies. The Operating Model also includes the Observation Error
Model (OEM), which generates simulated fishery data.

# Operating Model Conditioning

Conditioning is the process of developing an simulation or Operating
Model using the available data and knowledge. Operating Models should
represent plausible alternative hypotheses, and reflect the concerns of
stakeholders, since involving stakeholders is crucial to ensure that the
Management Strategy Evaluation process contributes to managing fisheries
in a way that respects both ecological and social aspects of
sustainability [@miller2019improving, @dankel2012advice,
@goethel2019recent]

A definition of a plausible scenario is one that fits prior knowledge,
with many sources of corroboration, without the complexity of
explanation, and with minimal conjecture [@connell2006model].
Plausibility may be determined formally, e.g. based on a statistical
approach to determine whether a system equivalent to the model generated
the data [@kell2021validation] or on expert judgement [e.g.,
@leach2014identification].

Stock assessments are commonly used to condition OMs, however there is
not always sufficient information to evaluate data quality and estimate
model parameters [@board1998improving]. A reason for this is because
observations are seldom complete and variables of interest are latent.
Therefore, there is often insufficient information to estimate key
parameters [@lee2012steepness,@lee2011m], and alternative choices may
provide equally good fits [@barlow_bartholomew1972]. For example, in an
age-based assessment, the production function and hence reference points
are largely determined by fixed parameters
e.g. natural mortality, and growth and fecundity-at-age, and steepness @mangel2013perspective.

The main sources of uncertainty in stock assessment are i) parameters
for which there is little information in the data and so are fixed (e.g.
natural mortality), ii) model structure, and iii) future process
variability, e.g. recruitment, [@PuntBM2016]. Therefore to ensure
robustness often requires conducting sensitivity tests or building
ensembles based on alternative assessment scenarios, for example using
hypotheses based on life history theory [@hamel2023natural]. Life
history theory has also been used tpo generate OMs for data limited
stocks see @fischer2020linking, @walker2023defining.

If alternative models explain the data equally well, it does not
necessarily follow that a model provides a true account of how the data
arose [@hodges1992you] Therefore to evaluate robustness, management
strategies must be tested under a range of hypothetical but plausible
scenarios. For example, does a harvest strategy only achieve management
objectives if natural mortality and the form of the stock recruitment
relationship are assumed to be known.

In stock assessment sensitivity and uncertainty analyses are often used
to understand the impacts of decisions in the face of uncertainty. In
sensitivity analysis, a limited set of scenarios, which include the most
important uncertainties, are considered, while in uncertainty analysis,
the objective is to quantify the probability of specific output values.
In both cases, procedures adopted for the selection, rejection and
weighting of scenarios will affect the robustness of advice. Therefore,
a cherry-picking approach for choosing model configurations based on the
implications for management advice should be avoided. For this reason
how the choice of OMs is important as this will determine the
performance of a management strategy.

Therefore when conditioning OMs a reference set and robustness sets are
normaly agreed in advance, before running the simulation trials.

### Reference Set

When conditioning Operating Models, it is common practice to condition a
reference set, i.e. a limited set of scenarios, which include the most
important uncertainties in the model structure, parameters, and data.
These comprise alternative scenarios which have both high
plausibility,reflect the concerns of stakeholders, and are likely to
major impacts on the performance of Management Strategies.

Instead of conditioning a number of OMs for the reference set, an
alternative is to implement a single scenario but with priors that
reflect uncertainty in parameters. Only one OMs is implemented because
multiple simulation runs are not comparable with the estimation of
posterior distributions in a Bayesian analysis [@chen2003assessing,
@berger1994overview]. A potential problem with such an approach is lack
of considering variations in process and measurement errors among
simulation runs.

### Robustness Set

A robustness set examines performance beyond the range of the Reference
Set. These are typically considered to be less likely than the reference
set Operating Models, and often focus on particularly challenging
circumstances with potentially negative consequences to be avoided. A
benefit of the reference set is that the value-of-information can be
evaluated, i.e. is there a benefit in reducing uncertainty by conducting
scientific study or collecting improved data.

### Redundancy

*MSE is a form of classification problem, used to identify the best
performing management strategy based on a set of performance metrics or
statistics [@fischer2021using]. When there are multiple OMs and metrics
these may overlap, resulting in imbalances, which may mean result in
inefficiency or bias. @maillo2020redundancy have provided a package of
metrics for big data classification problems. In particular,
Neighborhood Density and Decision Tree Progression, which study density
and accuracy progression. In addition, they have proposed a number of
basic metrics that can quickly characterise datasets.*

## Methods

Operating Models are generally developed using a stock assessment
paradigm. However, there are many processes for which there is little
information for in stock assessment datasets, e.g. natural mortality,
the steepness of the stock-recruitment relationship, the form of
population regulation and density dependence. Therefore, life history
relationships were used to create Operating Models based on life-history
relationships. This allows operating models to be conditioned on
ecological processes, of particular value in data-poor situations where
knowledge and data are limited, but also in data-rich situations as
simulation testing an assessment procedure using a model conditioned on
the same assumptions is not necessarily a true test of robustness.

To create an Operating Model, the FLR [@kell2007flr] packages mydas and
FLife were used. First an equilibrium per-recruit model was
parameterised for growth, maturity and natural mortality-at-age. The
per-recruit model then was combined with a stock recruit relationship
[@beverton1993dynamics].

A reference set was conditioned using a multivate distribution for x,y,
... to model uncertainty about parameters. While a reference set was
conditioined to represent uncetainty about the form of processes. Virgin
biomass was set at a constant value across all stocks and scenarios, so
that results are comparable, i.e presented in terms of exploitation
level and relative stock size and catch.

Life history parameters for growth, natural mortality and maturity were
used to develop an age-based Operating Model by parameterising
functional forms for weight ($W$), proportion mature ($Q$), natural
mortality ($M$) and fishing mortality ($F$) at age. These were then used
to calculate spawner ($S/R$) and yield-per-recruit ($Y/R$), which when
combined with a stock-recruitment relationship
[@sissenwine1987alternative] can be used to calculate the equilibrium
stock size as a function of fishing mortality ($F$). The equilibrium
relationships can then be turned into a dynamic model and projected
forward.

An equilibrium per-recruit model was parameterised for growth, maturity
and natural mortality-at-age. The per-recruit model was combined with a
stock-recruit relationship [@beverton1993dynamics]. An age-structured
simulation model was conditioned using life history theory to provide a
theoretical basis for developing hypotheses about population dynamics.
The parameters were: growth model parameters [$k$, $L_{\infty}$, $t_0$,
@vonbert1957quantitative]; $a$ and $b$ of the length-weight
relationship; and the length at which 50% were mature ($L_{50}$).
Natural mortality-at-age was modelled as a function of length
[@gislason2008does] and spawning stock biomass (SSB) was used as a proxy
for stock reproductive potential [SRP, @trippel_estimation_1999]. It was
assumed that fecundity is proportional to the weight-at-age of the
sexually mature portion of the population irrespective of the
demographic composition of adults [@murawski_impacts_2001] and that
processes such as sexual maturity are simple functions of age and
independent of sex [@matsuda_inconsistency_1996].

A variety of functional forms can be assumed for growth, natural
mortality, maturity, the selection pattern of the fisheries and the
stock-recruitment relationship. Processes such as growth and
maturity-at-age are generally observable, and data can be collected to
identify the correct functional form and to estimate parameters. The
functional form and parameters for processes such as natural mortality
and the stock-recruitment relationship are typically poorly known
[@michielsens2004bayesian]. Therefore, assumptions have to be made and
to evaluate the sensitivity of any analysis to those assumptions a
variety of scenarios need to be considered.

<!-- Life history parameters were extracted from FishBase (www.fishbase.org), -->

<!-- and Figure \ref{fig:lh} summarises $L_{\infty}$, $k$, $L_{50}$ and $b$; -->

<!-- with species ordered by $k$. There are only a few observations for -->

<!-- pollack and brill, which implies high uncertainty, since data in -->

<!-- FishBase are often based on small sample sizes, have limited coverage, -->

<!-- and life history parameters (e.g. maturity and growth) generally come -->

<!-- from different studies. There are clear relationships, both between and -->

<!-- within species, as $L_{\infty}$ is inversely correlated with $k$, and -->

<!-- $L_{50}$ is correlated with $L_{\infty}$. Brill has a value of $k$ -->

<!-- similar to sprat but its value of $L_{\infty}$ is closer to that of -->

<!-- turbot, while sprat has a large variation in the relationship between -->

<!-- length and weight (i.e. $b$ in the length-weight relationship $W=aL^b$). -->

Historical exploitation was simulated where

<!-- stocks were initially -->

<!-- lightly exploited before fishing mortality ($F$) gradually increased -->

<!-- until the stock became over-fished, after which a recovery plan was -->

<!-- implemented to bring fishing down to 70% of $F_{MSY}$ (Figure -->

<!-- \ref{fig:om}). This exploitation history provides contrasting periods of -->

<!-- under-, over- and maximally sustainable exploitation.  -->

### Scenarios

Even for data-rich stock assessments there is often large uncertainty
about the dynamics i.e. model uncertainty, [@punt2008refocusing], for
example estimates of stock status are highly sensitive to assumptions
about natural mortality-at-age [@jiao2012modelling], vulnerability of
age classes to the fisheries [@brooks2009analytical], and the
relationship between stock and recruitment [@cury2014resolving].
Therefore, a reference case was defined and a robustness set developed
representing the main sources of uncertainty (Table \ref{tab:om});
<!-- namely the -->
<!-- steepness of the stock recruitment relationship, recruitment -->
<!-- variability, natural mortality, selection pattern, and sample size -->
<!-- [@ono2015importance,@boorman1997recognising]. -->

Selectivity depends on the vulnerability of individuals to fishing, and
is typically either asymptotic or dome-shaped. The former indicates an
initial increase with age or size followed by a levelling off, while in
the later case selectivity-at-age declines. Selectivity will differ
between fisheries depending on gear characteristics, when and where the
fishery operates and the biology of the species. For example, estimates
for species captured with gill nets or hooks will be biased if the model
incorrectly assumes logistic selectivity because the missing large fish
will be assumed to have been caught, whereas they could just be missed
by the gear. Fisheries selectivity was therefore modelled as a double
normal, as this allowed both asymptotic and dome shaped selectivity to
be simulated. Logistic selectivity in the base case was based on the
maturity ogive, so that $MSY$ reference points are comparable across
case studies. The consequences of shifting the selection patterns is
well known, since if you fish below or above $L_{opt}$ (the length at
which a cohort attains its maximum biomass) you reduce $_{MSY}$, while
if you take fish before $L_{50}$ $F_{MSY}$ is reduced and $B_{MSY}$ is
increased as older fish need to be conserved. Therefore, scenarios were
not case-specific

### Life history processes

#### Individual Growth

Growth in length in this example was assumed to follow the Von
Bertalanffy growth equation @vonbert1957quantitative

```{=tex}
\begin{equation} L = L_\infty(1 - exp(-k(t-t_0)) \end{equation}
```
where $k$ is the rate at which the rate of growth in length declines as
length approaches the asymptotic length $L_\infty$ and $t_{0}$ is the
hypothetical time at which an individual is of zero length.

Length is converted to mass using the length-weight relationship

```{=tex}
\begin{equation} W = aL_t^b \end{equation}
```
\noindent where $a$ is the condition factor and $b$ is the allometric
growth coefficient.

#### Maturity-at-age

Proportion mature-at-age is modelled by the logistic equation with 3
parameters: age at 50% ($a_{50}$), 95% ($a_{95}$) mature, and maturity
as age approaches infinity.

```{=tex}
\begin{equation}
f(x) = \left\{ \begin{array}{ll}
            0                                 &\mbox{ if $(a_{50}-x)/a_{95} >  5$} \\
            a_{\infty}                        &\mbox{ if $(a_{50}-x)/a_{95} < -5$} \\
            \frac{m_{\infty}}{1.0+19.0^{(a_{50}-x)/_{95})}} &\mbox{ otherwise}
        \end{array}
       \right.
\end{equation}
```
#### Natural Mortality

The natural mortality of exploited fish populations is often assumed to
be a species-specific constant independent of body size. This assumption
has important implications for size-based fish population models and for
predicting the outcome of size-dependent fisheries management measures
such as mesh-size regulations [@gislason2008does]. Direct estimates of
the instantaneous natural mortality made in controlled studies, however,
are shown to vary by age [@lorenzen2002density]. Although $M$ can
sometimes be estimated within an assessment model, for example where
data from tagging provide information independent of fishing mortality
rates. However, in most cases, $M$ is derived from life-history
relationships, e.g. based on size
[@jensen1985comparison,@griffiths2007natural]. The large and
ever-increasing literature on this subject is a reflection of
uncertainty. @gislason2008does in an empirical study showed that $M$ is
significantly related to body length, asymptotic length and k.
Temperature is non-significant when $k$ is included, since $k$ itself is
correlated with temperature, i.e.

```{=tex}
\begin{equation}
            M =  0.55L^{1.61}L_\infty^{1.44}k
\end{equation}
```
#### Selection Pattern

In this study, the vulnerability of individuals to fishing was modelled
by the selection pattern. Typically, selectivity is either asymptotic or
dome-shaped. The former indicates an initial increase with age or size
followed by a levelling off, while in the latter case selectivity-at-age
declines. Selectivity will differ between fisheries depending on gear
characteristics, when and where the fishery operates and the biology of
the species. Fisheries selectivity was therefore modelled as a double
normal, as this allowed both asymptotic and dome-shaped selectivity to
be simulated. The double normal has three parameters that describe the
age at maximum selection ($a_1$), the rate at which the left-hand limb
increases ($sl$) and the right-hand limb decreases ($sr$) which allows
flat topped or domed shaped selection patterns to be chosen, i.e.

```{=tex}
\begin{equation}
f(x) = \left\{ \begin{array}{ll}
            2^-(\frac{a-a_1}{s_l})^2 , & if a < a_1        \\
            2^-(\frac{a-a_1}{s_r})^2 , & if a \geq a_1      \\
            \end{array}
       \right.
\end{equation}
```
#### Spawning Stock Biomass

Spawning stock biomass ($S$), is the sum of the products of the numbers
of females, $N$, proportion mature-at-age, $Q$ and their mean
fecundity-at-age, $G$, which is taken to be proportional to their
weight-at-age i.e.

```{=tex}
\begin{equation} S = \sum\limits_{i=0}^{p} {N_iQ_iG_i} \end{equation}
```
In this study, fecundity-at-age is assumed proportional to biomass and
the sex ratio to be 1:1.

#### Stock Recruitment Relationship

A Beverton and Holt stock-recruitment relationship @beverton1993dynamics
was assumed. This relationship is derived from a simple
density-dependent mortality model where the more survivors there are the
higher the mortality. It is assumed that the number of recruits ($R$)
increases towards an asymptotic level ($R_{max}$) as $S$ increases, i.e.

```{=tex}
\begin{equation} R=Sa/(b+S) \end{equation}
```
The relationship was reformulated in terms of steepness ($h$), virgin
biomass ($v$) and $S/R_{F=0}$. Where steepness is the proportion of the
expected recruitment produced at 20% of virgin biomass relative to
virgin recruitment $(R_0)$. However, there is often insufficient
information to allow its estimation from stock assessment
[@pepin2015reconsidering] and so scenarios were considered for different
value of steepness, i.e. resilience to overfishing. Virgin biomass was
set at 1000 Mt to allow comparisons to be made across scenarios.

```{=tex}
\begin{equation}
R=\frac{0.8R_0h}{0.2S/R_{F=0}R_0(1-h)+(h-0.2)S}
\end{equation}
```
### Equilibrium Analysis

@sissenwine1987alternative, estimated surplus production using an
age-based analysis using an equilibrium analysis by combining a
stock-recruitment relationship, a spawning-stock-biomass-per-recruit
analysis, and a yield-per-recruit analysis. For any specified rate of
fishing mortality, an associated value of spawning stock biomass ($S$)
per recruit ($R$) is $S/R$ is defined, based on the assumed processes
for growth, natural mortality and selection pattern-at-age detailed in
the previous sections.

```{=tex}
\begin{equation}
S/R=\sum\limits_{i=0}^{p-1} {e^{\sum\limits_{j=0}^{i-1} {-F_j-M_j}}} W_i Q_i + e^{\sum\limits_{i=0}^{p-1} {-F_i-M_i}} \frac{W_p Q_p}{1-e^{-F_p-M_p}}\\
\end{equation}
```
When the value of $S/R$ obtained is inverted and superimposed on the
stock-recruitment function as a slope ($R/S$), the intersection of this
slope with the stock-recruitment function defines an equilibrium level
of recruitment. When this value of recruitment is multiplied by the
yield per recruit calculated for the same fishing mortality rate, the
equilibrium yield associated with the fishing mortality rate emerges
[@gabriel1999review].

```{=tex}
\begin{equation}
Y/R=\sum\limits_{a=r}^{n-1} {e^{\sum\limits_{i=r}^{a-1} {-F_i-M_i}}} W_a\frac{F_a}{F_a+M_a}\left(1-e^{-F_i-M_i} \right) + e^{\sum\limits_{i=r}^{n-1} {-F_n-M_n}} W_n\frac{F_n}{F_n+M_n}\\
\end{equation}
```
The second term is the plus-group, i.e. the summation of all ages from
the last age to infinity.

### Projection

The stock-recruitment relationship and the vectors of weight, natural
mortality, maturity and selectivity-at-age allow a forward projection
model to be created, which forms the basis of the Operating Model.

```{=tex}
\begin{equation}
N_{t,a} =
\left\{
\begin{array}{ll}
\vspace{0.2cm}
R_t, & \rm{if} \ a=0,\\
\vspace{0.2cm}
N_{t-1,a-1} e^{-Z_{t-1,a-1}}, & \rm{if} \ 1 \leq a \leq A-1,\\
N_{t-1,A-1} e^{-Z_{t-1,A-1}}+N_{t-1,A} e^{-Z_{t-1,A}}, & \rm{if} \ a=A,
\end{array}
\right.
\end{equation}
```
Where $N_{t,a}$ is the number of fish of age $a$ at the beginning of
year $t$, $R_t$ is the total number of recruits born in year $t$. Here,
$A$ is the so-called plus group age, which is an aggregated age greater
than or equal to the actual age $A$.

### Seasonal Operating Model

A seasonal Operating Model was developed to explore the impact of
shorter lags between survey, advice and exploitation. This allowed
alternative management options such as in-season rules to be considered.

### Stock Components or Mixed Stocks

## Software Versions

-   `r version$version.string`

-   FLCore: `r packageVersion('FLCore')`

-   FLPKG: `r # packageVersion('FLPKG')`

-   FLBRP: `r packageVersion('FLBRP')`

-   FLasher: `r packageVersion('FLasher')`

-   ggplotFL: `r packageVersion('ggplotFL')`

-   FLife: `r packageVersion('FLife')`

-   mydas: `r packageVersion('mydas')`

-   FLCandy: `r packageVersion('FLCandy')`

-   **Compiled**: `r date()`

-   **Git Hash**:
    `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence KELL**.
[laurie\@seaplusplus.co.uk](mailto:laurie@seaplusplus.co.uk){.email}

# References
